---
title: "Smoking Cessation: Pre-Processing"
author: "Zoë E. Laky"
date: "20250530"
output: html_notebook
---

## Installing & Loading Packages (Part 1)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/LV")

list.of.packages <- c('tidyverse', 'dplyr', 'ggplot2', 'caret', 'visdat', 'moments', 'arsenal', 'sjPlot', 'sjmisc', 'parameters', 'DataExplorer')
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) 
  install.packages(new.packages)

packages <- c('tidyverse', 'dplyr', 'ggplot2', 'caret', 'visdat', 'moments', 'arsenal', 'sjPlot', 'sjmisc', 'parameters', 'DataExplorer')   
lapply(packages, require, character.only = TRUE)

citation("tidyverse")
citation("dplyr")
citation("ggplot2")
citation("caret")
citation("visdat")
citation("moments")
citation("arsenal")
citation("sjPlot")
citation("sjmisc")
citation("parameters")
citation("DataExplorer")

getwd()
sessionInfo()

rm(list.of.packages, new.packages, packages)
```

## Loading Data (Part 1)
```{r}
cessationP1 <- read.csv('LVDATA_FULL_8.11.19.csv', header = T, na.strings = "", stringsAsFactors = T) #importing data

head(cessationP1) #view first few rows of data
dim(cessationP1) #dimensions of dataset
table(is.na(cessationP1)) #total missing values ("NA") in dataset 
sapply(cessationP1, function(x) sum(is.na(x))) #total missing values ("NA") in each column
```

## Dropout & Qualtrics Errors ID List
```{r}
#Dropout
dropped <- cessationP1[ , c("ID", "Dropout")]

droppedy <- dropped[ which(dropped$Dropout==1), ] #select cases in dropped for which Dropout = 1 (Dropped out/lost to follow-up)
table(droppedy$Dropout)
write.csv(droppedy, "~/Desktop/LV/droppedy.csv", row.names=FALSE)

droppedn <- dropped[ which(dropped$Dropout==0), ] #select cases in dropped for which Dropout = 0 (Completed study)
table(droppedn$Dropout)
write.csv(droppedn, "~/Desktop/LV/droppedn.csv", row.names=FALSE)

#Qualtrics Errors
#Aims(s):
  #Create a dummy variable for each question set (0 if missing all pieces, 1 if any of the questions within the set are answered). Sum columns to get the total question sets answered. Expect 7 or 9 (exclude dropouts).
  #Compare to list created by study visit dates (correction on June 24, 2016; 1 = affect by Qualtrics error)

cessationP1$AdminIssuef <- NA

cessationP1$AdminIssuef = ifelse(cessationP1$ID %in%
  c("LV001", "LV148", "LV221", "LV075", "LV147", "LV150", "LV223", "LV002", "LV224", "LV152", "LV153", "LV154", "LV076", "LV155", "LV003", "LV004", "LV005", "LV157", "LV226", "LV006", "LV227", "LV225", "LV158", "LV156", "LV077", "LV159", "LV160", "LV163", "LV162", "LV228", "LV007", "LV164", "LV230", "LV229", "LV166", "LV079", "LV167", "LV168", "LV231", "LV169", "LV080", "LV232","LV171", "LV172", "LV081", "LV173", "LV009", "LV174", "LV233", "LV083", "LV234", "LV085", "LV084", "LV082"), 1, 0) #create list of first 54 participants who completed follow-up visit before correction on June 24, 2016 (participation log referenced)
table(cessationP1$AdminIssuef)

adminissue <- cessationP1[ , c("ID", "AdminIssuef")]

adminissuey <- adminissue[ which(adminissue$AdminIssuef==1), ] #select cases for which AdminIssuef = 1 (Administration Issue)
table(adminissuey$AdminIssuef)
write.csv(adminissuey, "~/Desktop/LV/adminissuey.csv", row.names=FALSE)

adminissuen <- adminissue[ which(adminissue$AdminIssuef==0), ] #select cases for which AdminIssuef = 0 (No Administration Issue)
table(adminissuen$AdminIssuef)
write.csv(adminissuen, "~/Desktop/LV/adminissuen.csv", row.names=FALSE)

cessationP1$FTNDfMiss <- 1 #create new variable for FTND follow-up block where all questions were missed 
cessationP1$FTNDfMiss[is.na(cessationP1$FTND1_FU) & is.na(cessationP1$FTND2_FU) & is.na(cessationP1$FTND3_FU) & is.na(cessationP1$FTND4_FU) & is.na(cessationP1$FTND5_FU) & is.na(cessationP1$FTND6_FU)] <- 0

cessationP1$CLfMiss <- 1 #create new variable for CL follow-up block where all questions were missed 
cessationP1$CLfMiss[is.na(cessationP1$CL2_FU)] <- 0

cessationP1$SoCfMiss <- 1 #create new variable for SoC follow-up block where all questions were missed 
cessationP1$SoCfMiss[is.na(cessationP1$SoC1_FU) & is.na(cessationP1$SoC2_FU) & is.na(cessationP1$SoC5_FU)] <- 0

cessationP1$SSEQfMiss <- 1 #create new variable for SSEQ follow-up block where all questions were missed 
cessationP1$SSEQfMiss[is.na(cessationP1$SSEQ1_FU) & is.na(cessationP1$SSEQ2_1_FU) & is.na(cessationP1$SSEQ3_FU) & is.na(cessationP1$SSEQ4_1_FU) & is.na(cessationP1$SSEQ5_FU) & is.na(cessationP1$SSEQ6_1_FU) & is.na(cessationP1$SSEQ7_FU) & is.na(cessationP1$SSEQ8_1_FU) & is.na(cessationP1$SSEQ9_FU) & is.na(cessationP1$SSEQ10_1_FU) & is.na(cessationP1$SSEQ11_FU) & is.na(cessationP1$SSEQ12_1_FU) & is.na(cessationP1$SSEQ13_FU) & is.na(cessationP1$SSEQ14_1_FU) & is.na(cessationP1$SSEQ15_FU) & is.na(cessationP1$SSEQ16_1_FU) & is.na(cessationP1$SSEQ17_FU) & is.na(cessationP1$SSEQ18_1_FU) & is.na(cessationP1$SSEQ19_FU) & is.na(cessationP1$SSEQ20_1_FU) & is.na(cessationP1$SSEQ21_FU) & is.na(cessationP1$SSEQ22_1_FU) & is.na(cessationP1$SSEQ23_FU) & is.na(cessationP1$SSEQ24_1_FU) & is.na(cessationP1$SSEQ25_FU) & is.na(cessationP1$SSEQ26_1_FU) & is.na(cessationP1$SSEQ27_FU) & is.na(cessationP1$SSEQ28_1_FU) & is.na(cessationP1$SSEQ29_FU) & is.na(cessationP1$SSEQ30_1_FU) & is.na(cessationP1$SSEQ31_FU) & is.na(cessationP1$SSEQ32_1_FU) & is.na(cessationP1$SSEQ33_FU) & is.na(cessationP1$SSEQ34_1_FU)] <- 0

cessationP1$SCQfMiss <- 1 #create new variable for SCQ follow-up block where all questions were missed 
cessationP1$SCQfMiss[is.na(cessationP1$SCQ1_FU) & is.na(cessationP1$SCQ2_FU) & is.na(cessationP1$SCQ3_FU) & is.na(cessationP1$SCQ4_FU) & is.na(cessationP1$SCQ5_FU) & is.na(cessationP1$SCQ6_FU) & is.na(cessationP1$SCQ7_FU) & is.na(cessationP1$SCQ8_FU) & is.na(cessationP1$SCQ9_FU) & is.na(cessationP1$SCQ10_FU) & is.na(cessationP1$SCQ11_FU) & is.na(cessationP1$SCQ12_FU) & is.na(cessationP1$SCQ13_FU) & is.na(cessationP1$SCQ14_FU) & is.na(cessationP1$SCQ15_FU) & is.na(cessationP1$SCQ16_FU) & is.na(cessationP1$SCQ17_FU) & is.na(cessationP1$SCQ18_FU) & is.na(cessationP1$SCQ19_FU) & is.na(cessationP1$SCQ20_FU) & is.na(cessationP1$SCQ21_FU) & is.na(cessationP1$SCQ22_FU) & is.na(cessationP1$SCQ23_FU) & is.na(cessationP1$SCQ24_FU) & is.na(cessationP1$SCQ25_FU) & is.na(cessationP1$SCQ26_FU) & is.na(cessationP1$SCQ27_FU) & is.na(cessationP1$SCQ28_FU) & is.na(cessationP1$SCQ29_FU) & is.na(cessationP1$SCQ30_FU) & is.na(cessationP1$SCQ31_FU) & is.na(cessationP1$SCQ32_FU) & is.na(cessationP1$SCQ33_FU) & is.na(cessationP1$SCQ34_FU) & is.na(cessationP1$SCQ35_FU) & is.na(cessationP1$SCQ36_FU) & is.na(cessationP1$SCQ37_FU) & is.na(cessationP1$SCQ38_FU) & is.na(cessationP1$SCQ39_FU) & is.na(cessationP1$SCQ40_FU) & is.na(cessationP1$SCQ41_FU) & is.na(cessationP1$SCQ42_FU) & is.na(cessationP1$SCQ43_FU) & is.na(cessationP1$SCQ44_FU) & is.na(cessationP1$SCQ45_FU) & is.na(cessationP1$SCQ46_FU) & is.na(cessationP1$SCQ47_FU) & is.na(cessationP1$SCQ48_FU) & is.na(cessationP1$SCQ49_FU) & is.na(cessationP1$SCQ50_FU) & is.na(cessationP1$SCQ51_FU) & is.na(cessationP1$SCQ52_FU) & is.na(cessationP1$SCQ53_FU) & is.na(cessationP1$SCQ54_FU) & is.na(cessationP1$SCQ55_FU) & is.na(cessationP1$SCQ_OE_FU)] <- 0

cessationP1$PRBQfMiss <- 1 #create new variable for PRBQ follow-up block where all questions were missed 
cessationP1$PRBQfMiss[is.na(cessationP1$PRBQ1_FU) & is.na(cessationP1$PRBQ2_FU) & is.na(cessationP1$PRBQ3_FU) & is.na(cessationP1$PRBQ4_FU) & is.na(cessationP1$PRBQ5_FU) & is.na(cessationP1$PRBQ6_FU) & is.na(cessationP1$PRBQ7_FU) & is.na(cessationP1$PRBQ8_FU) & is.na(cessationP1$PRBQ9_FU) & is.na(cessationP1$PRBQ10_FU) & is.na(cessationP1$PRBQ11_FU) & is.na(cessationP1$PRBQ12_FU) & is.na(cessationP1$PRBQ13_FU) & is.na(cessationP1$PRBQ14_FU) & is.na(cessationP1$PRBQ15_FU) & is.na(cessationP1$PRBQ16_FU) & is.na(cessationP1$PRBQ17_FU) & is.na(cessationP1$PRBQ18_FU) & is.na(cessationP1$PRBQ19_FU) & is.na(cessationP1$PRBQ20_FU) & is.na(cessationP1$PRBQ21_FU) & is.na(cessationP1$PRBQ22_FU) & is.na(cessationP1$PRBQ23_FU) & is.na(cessationP1$PRBQ24_FU) & is.na(cessationP1$PRBQ25_FU) & is.na(cessationP1$PRBQ26_FU) & is.na(cessationP1$PRBQ27_FU) & is.na(cessationP1$PRBQ28_FU) & is.na(cessationP1$PRBQ29_FU) & is.na(cessationP1$PRBQ30_FU) & is.na(cessationP1$PRBQ31_FU) & is.na(cessationP1$PRBQ32_FU) & is.na(cessationP1$PRBQ33_FU) & is.na(cessationP1$PRBQ34_FU) & is.na(cessationP1$PRBQ35_FU) & is.na(cessationP1$PRBQ36_FU) & is.na(cessationP1$PRBQ37_FU) & is.na(cessationP1$PRBQ38_FU) & is.na(cessationP1$PRBQ39_FU) & is.na(cessationP1$PRBQ40_FU) & is.na(cessationP1$PRBQ_OE_FU)] <- 0

cessationP1$SGOfMiss <- 1 #create new variable for SGO follow-up block where all questions were missed 
cessationP1$SGOfMiss[is.na(cessationP1$SGOFU_Coded) & is.na(cessationP1$SGO2_FU) & is.na(cessationP1$SGO3_FU)] <- 0

cessationP1$QuitfMiss <- 1 #create new variable for Quit follow-up block where all questions were missed 
cessationP1$QuitfMiss[is.na(cessationP1$QuA1)] <- 0
#Note:
  #Of those that missed QuA1, 13 participants have response of "0" for QuA2. Unclear how this is possible and will be considered a data entry error in QuA2.

cessationP1$OSfMiss <- 1 #create new variable for OS follow-up block where all questions were missed 
cessationP1$OSfMiss[is.na(cessationP1$OtherStudiesYN)] <- 0

cessationP1$SysMissf <- cessationP1$FTNDfMiss + cessationP1$CLfMiss + cessationP1$SoCfMiss + cessationP1$SSEQfMiss + cessationP1$SCQfMiss + cessationP1$PRBQfMiss + cessationP1$SGOfMiss + cessationP1$QuitfMiss + cessationP1$OSfMiss
table(cessationP1$SysMissf)

sysmissf <- cessationP1[ , c("ID", "Dropout", "AdminIssuef", "SysMissf", "FTNDfMiss", "CLfMiss", "SoCfMiss", "SSEQfMiss", "SCQfMiss", "PRBQfMiss", "SGOfMiss", "QuitfMiss", "OSfMiss")]
table(sysmissf$Dropout, cessationP1$SysMissf)

sysmissff1 <- sysmissf[which(sysmissf$AdminIssuef==1 & sysmissf$Dropout==0), ]

sysmissff2 <- cessationP1[which(cessationP1$ID=="LV227"), ]
#Note: 
  #LV227 - Responses indicate participant saw 8 of the 9 questionnaires despite being within the first 54 people to complete the follow-up visit. Most likely case is that participant did not see QuA1 and QuA2 (responses of 0 may be data entry errors when they should be "NA")

cessationP1$QuitfMissRc <- cessationP1$QuitfMiss
cessationP1$QuitfMissRc[cessationP1$ID == "LV227" & cessationP1$QuitfMiss==1] <- 0

sysmissff3 <- sysmissf[which(sysmissf$AdminIssuef==0 & sysmissf$Dropout==0), ]
table(sysmissff3$QuitfMiss) #participants who did not drop out and saw QuA1, but did not respond to QuA1

rm(adminissue, adminissuen, adminissuey, dropped, droppedn, droppedy, sysmissf, sysmissff1, sysmissff2, sysmissff3)
```

## Remove Variables (Part 1)
```{r}
cessationP2 <- subset(cessationP1, select = -c(FTNDfMiss, CLfMiss, SoCfMiss, SSEQfMiss, SCQfMiss, PRBQfMiss, SGOfMiss, OSfMiss, QuitfMiss, avg.cigsday_FU, Brand_FU, cigsinpast24hours_FU, COReading_FU, CL2_FU, ContemplationStatus_FU, ecigpastmonth_FU, Flavor_FU, FTND1_FU, FTND2_FU, FTND3_FU, FTND4_FU, FTND5_FU, FTND6_FU, Menthol_FU, OtherStudiesYN, OtherStudies., PRBQ10_FU, PRBQ11_FU, PRBQ12_FU, PRBQ13_FU, PRBQ14_FU, PRBQ15_FU, PRBQ16_FU, PRBQ17_FU, PRBQ18_FU, PRBQ19_FU, PRBQ1_FU, PRBQ20_FU, PRBQ21_FU, PRBQ22_FU, PRBQ23_FU, PRBQ24_FU, PRBQ25_FU, PRBQ26_FU, PRBQ27_FU, PRBQ28_FU, PRBQ29_FU, PRBQ2_FU, PRBQ30_FU, PRBQ31_FU, PRBQ32_FU, PRBQ33_FU, PRBQ34_FU, PRBQ35_FU, PRBQ36_FU, PRBQ37_FU, PRBQ38_FU, PRBQ39_FU, PRBQ3_FU, PRBQ40_FU, PRBQ4_FU, PRBQ5_FU, PRBQ6_FU, PRBQ7_FU, PRBQ8_FU, PRBQ9_FU, PRBQOE_FU_Coded, SCQ10_FU, SCQ11_FU, SCQ12_FU, SCQ13_FU, SCQ14_FU, SCQ15_FU, SCQ16_FU, SCQ17_FU, SCQ18_FU, SCQ19_FU, SCQ1_FU, SCQ20_FU,SCQ21_FU, SCQ22_FU, SCQ23_FU, SCQ24_FU, SCQ25_FU, SCQ26_FU, SCQ27_FU, SCQ28_FU, SCQ29_FU,SCQ2_FU, SCQ30_FU, SCQ31_FU, SCQ32_FU, SCQ33_FU, SCQ34_FU, SCQ35_FU, SCQ36_FU, SCQ37_FU, SCQ38_FU, SCQ39_FU, SCQ3_FU, SCQ40_FU, SCQ41_FU, SCQ42_FU, SCQ43_FU, SCQ44_FU, SCQ45_FU, SCQ46_FU, SCQ47_FU, SCQ48_FU, SCQ49_FU, SCQ4_FU, SCQ50_FU, SCQ51_FU, SCQ52_FU, SCQ53_FU, SCQ54_FU, SCQ55_FU, SCQ5_FU, SCQ6_FU, SCQ7_FU, SCQ8_FU, SCQ9_FU, SCQOE_FU_Coded, SGO2_FU, SGO3_FU, SGOFU_Coded, Smokedmarijuana_FU, SoC1_FU, SoC2_FU, SoC5_FU, VAS_2, FTND_Total_FU, Abstinence_FU, Amtmarijuanasmoked_FU, brand_ecig_FU, DateofBirth, Ethnicity_Missing0, hourssincelastcig_FU, menthol_ecig_FU, Race_Missing0, SCQ_OE_FU, SGO1_FU, Smokedinpast24hours_FU, SoC3_FU, SoC4_FU, SoC6_FU, SoC7_FU, SSEQ_Answered_FU, SSEQ2_1_FU, SSEQ4_1_FU, SSEQ6_1_FU, SSEQ8_1_FU, SSEQ10_1_FU, SSEQ12_1_FU, SSEQ14_1_FU, SSEQ16_1_FU, SSEQ18_1_FU,  SSEQ20_1_FU, SSEQ22_1_FU, SSEQ24_1_FU, SSEQ26_1_FU, SSEQ28_1_FU, SSEQ30_1_FU, SSEQ32_1_FU, SSEQ34_1_FU, SSEQ1_FU, SSEQ3_FU, SSEQ5_FU, SSEQ7_FU, SSEQ9_FU, SSEQ11_FU, SSEQ13_FU, SSEQ15_FU, SSEQ17_FU, SSEQ19_FU, SSEQ21_FU, SSEQ23_FU, SSEQ25_FU, SSEQ27_FU, SSEQ29_FU, SSEQ31_FU, SSEQ33_FU, type_ecig_FU,  PRBQ_Ben_FU,PRBQ_Risk_FU, PRBQ_OE_FU, SCQA_Ben_FU, SCQA_Risk_FU, SSEQ_Total_FU)) #removing follow-up variables (except primary outcome) 
names(cessationP2) 
write.csv(cessationP2, "~/Desktop/LV/cessationP2.csv", row.names=FALSE)

sapply(cessationP2, function(x) sum(is.na(x))) #total missing values ("NA") in each column
dim(cessationP2) #dimensions of data set
table(is.na(cessationP2)) #total missing values ("NA") in data set 
```

## Exploratory (Part 1): Dropout
```{r}
cessationP2$Dropout <- factor(cessationP2$Dropout)
table(cessationP2$Dropout)

plotcounts1 <- table(cessationP2$Dropout, cessationP2$looming)
mosaicplot(plotcounts1, xlab='Dropout Status', ylab='Condition', main='Dropout by Condition', col='blue')
chisq.test(cessationP2$Dropout, cessationP2$looming, correct=FALSE)

plotcounts2 <- table(cessationP2$Dropout, cessationP2$narrator)
mosaicplot(plotcounts2, xlab='Dropout Status', ylab='Narrator', main='Dropout by Narrator', col='green')
chisq.test(cessationP2$Dropout, cessationP2$narrator, correct=FALSE)

plotcounts3 <- table(cessationP2$Dropout, cessationP2$study_condition)
mosaicplot(plotcounts3, xlab='Dropout Status', ylab='Condition & Narrator', main='Dropout by Condition-Narrator', col='orange')

plotcounts4 <- table(cessationP2$Dropout, cessationP2$Gender)
mosaicplot(plotcounts4, xlab='Dropout Status', ylab='Gender', main='Dropout by Gender', col='pink')
chisq.test(cessationP2$Dropout, cessationP2$Gender, correct=FALSE)

plotcounts5 <- table(cessationP2$Dropout, cessationP2$Race)
mosaicplot(plotcounts5, xlab='Dropout Status', ylab='Race', main='Dropout by Race', col='purple')

t.test(cessationP2$Age~cessationP2$Dropout) 
t.test(cessationP2$CSCLSTotal~cessationP2$Dropout) 
t.test(cessationP2$FTND_Total~cessationP2$Dropout) 
t.test(cessationP2$PRBQ_Ben~cessationP2$Dropout) 
t.test(cessationP2$PRBQ_Risk~cessationP2$Dropout) 
t.test(cessationP2$SCQA_Ben~cessationP2$Dropout) 
t.test(cessationP2$SCQA_Risk~cessationP2$Dropout) 
t.test(cessationP2$SSSVTotalScore~cessationP2$Dropout) 
t.test(cessationP2$VVIQClosed~cessationP2$Dropout) 
t.test(cessationP2$VVIQOpen~cessationP2$Dropout) 

rm(plotcounts1, plotcounts2, plotcounts3, plotcounts4, plotcounts5)
```

## Remove Rows
```{r}
table(cessationP2$QuitfMissR, useNA="always")
table(cessationP2$Dropout, useNA="always")

cessationP3 <- cessationP2[which(cessationP1$Dropout==0 & cessationP1$QuitfMissRc==1), ] #excluding dropouts and missed primary outcome
write.csv(cessationP3, "~/Desktop/LV/cessationP3.csv", row.names=FALSE)
dim(cessationP3)
cessationP3 <- cessationP3[,order(colnames(cessationP3))] #sort variables alphabetically

table(cessationP3$Ethnicity, cessationP3$QuA1, useNA = 'always')
mean(cessationP3$Age, na.rm=TRUE)
sd(cessationP3$Age, na.rm=TRUE)

rm(cessationP1)
```

## Missing Data (Part 1)
```{r}
#Flag
  #Total: Missing(NA) = 10.8%
  #Row: Max = 35.779817% (118)
  #Column: Max = 100% (QuA6)

vis_miss(cessationP3)
(sum(is.na(cessationP3))/prod(dim(cessationP3)))*100 #percentage of missing data in entire data frame

colmiss <- (colMeans(is.na(cessationP3)))*100 #percentage of missing data by column
colmiss <- as.data.frame(colmiss)

mean(complete.cases(cessationP3))*100 #percentage of complete rows
rowmiss <- (rowSums(is.na(cessationP3)))/327*100
rowmiss <- as.data.frame(rowmiss)

rm(colmiss, rowmiss)
```

## Recode Variables (Part 1)

**Smoking**
```{r}
#Smoking History 
#Aim(s): 
  #Reduce dimensionality, while retaining as much information as possible
  #Collapse Cig8, Cig9, Cig10, and Cig11 into Cig9 (number of past quit attempts)
#Note:
  #Cig10, Cig11 - very difficult to clean (e.g., estimate)
  #See if the following IDs are miss-classified by the model: LV031, LV176, LV179, LV244, LV255, LV258
  #Cig9Rc and Previousquitattempt capture overlapping information, with Cig9Rc containing continuous data and Previousquitattempt=NA for 29 participants. Remove Previousquitattempt. 

smokehis <- cessationP3[ , c("ID", "Cig1", "Cig2", "Cig3", "Cig4", "Cig5", "Cig6", "Cig7", "Cig8", "Cig9", "Cig10", "Cig11", "Brand", "Flavor", "Menthol", "Smokedinpast24hours", "hourssincelastcig", "Previousquitattempt", "COReading", "cigsinpast24hours", "Avg.cigsday")] 

smokehisf1 <- smokehis[which(smokehis$Cig8==0), ] #select cases in smokehis for which Cig8 = 0 to check if anyone responded no to past quit attempts (Cig8), but then listed a number of past quit attempts (Cig9) 
table(smokehisf1$Cig9)
#Coding - no changes (Cig8 to be deleted)

smokehisf2 <- smokehis[which(smokehis$Cig8==1), ] #select cases in smokehis for which Cig8 = 1 to check if anyone responded yes to past quit attempts (Cig8), but then did not answer number of past quit attempts (Cig9) or reported quit attempt <24 hours
table(smokehisf2$Cig9, useNA = "always")

smokehisf3 <- smokehis[which(smokehis$Cig8==1 & is.na(smokehis$Cig9)), ] #select cases in smokehis for which Cig8 = 1, but Cig9 = "NA" to see cases where number of past quit attempts (Cig9) is not answered (IDs: LV065, LV353, LV378)
#Coding - Refer to level below: Cig10 and Cig11; Conditional that response is >24 hours
  #If Cig10 = Cig11 -> Cig9 = 1
  #If Cig10 = "NA" or (not both) Cig11 = "NA" -> Cig9 = 1
  #If Cig10 ≠ Cig11 -> Cig9 = 2 (if numbers differ then they have had at least 2 quit attempts so 2 is the best estimate)

smokehis$Cig9Rc <- smokehis$Cig9
smokehis$Cig9Rc[smokehis$ID == "LV065" & is.na(smokehis$Cig9)] <- 1
smokehis$Cig9Rc[smokehis$ID == "LV353" & is.na(smokehis$Cig9)] <- 2
smokehis$Cig9Rc[smokehis$ID == "LV378" & is.na(smokehis$Cig9)] <- 2

smokehisf4 <- smokehis[which(smokehis$Cig9==0), ] #selecting cases in smokehis for which Cig9=0 to see cases where number of past quit attempts (Cig9) is 0, but Cig10 and Cig11 > 0 (IDs: LV031, LV176, LV179, LV244, LV255, LV258)
#Coding - Do not change. Participants may not have been clear on skip logic and when they should restart answering questions. Given they responded to 2 questions that they have not made a quit attempt in the past, assume Cig10 and Cig11 were interpreted incorrectly.

table(smokehis$Cig8, useNA = "always") #Checking number of Cig8="NA" to determine if Cig9="NA" means not answered (i.e., missing data) or not applicable (i.e., followed appropriate skip logic)
table(smokehis$Cig9Rc, useNA = "always") 
#Note: Given no missing data for Cig8, all Cig9Rc="NA" were skipped as intended and should mean that there were 0 quit attempts. Check assumption by reviewing Cig10 and Cig11.

smokehisf5 <- smokehis[which(smokehis$Cig8==0 & is.na(smokehis$Cig9)), ] #selecting cases in smokehis for which Cig9=NA and Cig8=0 to see if any have responses in Cig10 and Cig11 
#Coding - Change Cig9="NA" to 0. Participants appropriately followed skip logic for Cig8 and Cig9. Participants may not have been clear on skip logic and when they should restart answering questions. Assume Cig10 and Cig11 were interpreted incorrectly. 

smokehis$Cig9Rc[smokehis$Cig8==0 & is.na(smokehis$Cig9)] <- 0
table(smokehis$Cig9Rc, useNA = "always") 
#Coding - Given no missing data for Cig8, if Cig9Rc="NA", recode as 0

table(smokehis$Brand) #frequencies of categorical responses
#Coding - American Spirit=1, 555's=2, Benson&Hedges=3, Camel=4, Cheyennes=5, Kool=6, L&M=7, Marlboro=8, Maverick=9, Newport=10, Pall Mall=11, Parliament=12, Pyramid=13, Salem=14, Self-Rolled=15

smokehis <- smokehis %>%
  mutate(BrandBRc = dplyr::case_when(
    Brand %in% c("American Spirit", "American Spirit Blue") ~ 1,
    Brand == "555's" ~ 2,
    Brand == "Benson & Hedges 100s" ~ 3,
    Brand %in% c("Camel", "Camel Crush", "Camel Filter") ~ 4,
    Brand == "Cheyennes" ~ 5,
    Brand %in% c("Cool", "Cool 100", "Kool", "Kool 100", "Kool Filter King", "Kools 100", "Kools Filter King") ~ 6,
    Brand == "L&M 100" ~ 7,
    Brand %in% c("Malboro", "Marlboro", "Marlboro 100", "Marlboro 100 Black", "Marlboro 100 Reds", 
                 "Marlboro 100 Special Ble", "Marlboro 27", "Marlboro Black 100", "Marlboro Blend No 27", 
                 "Marlboro Light", "Marlboro Reds", "Marlboro Silver") ~ 8,
    Brand %in% c("Maverick", "Maverick 100", "Maverick 100 and Newport", "Maverick Gold", 
                 "Maverick Green 100", "Maverick Menthol", "Maverick or Newport", "Mavericks (and Newport)") ~ 9,
    Brand %in% c("Newport", "Newport 100", "Newport and Kool", "Newport Medium", 
                 "Newport Red", "Newport Reds") ~ 10,
    Brand %in% c("Pall Mall", "Pall Mall Red 100", "PallMall") ~ 11,
    Brand == "Parliament" ~ 12,
    Brand == "Pyramid or Maverick red" ~ 13,
    Brand == "Salem or Newport" ~ 14,
    TRUE ~ NA_real_  # default if unmatched
  ))

table(smokehis$Flavor)
#Coding - Bold=1, Full Flavor=2, Regular=3, Light=4, Ultra Light=5
smokehis <- smokehis %>%
  mutate(FlavorBRc = dplyr::case_when(
    Flavor == "Bold" ~ 1,
    Flavor %in% c("Full flavor", "Full Flavor") ~ 2,
    Flavor == "Regular" ~ 3,
    Flavor == "Light" ~ 4,
    Flavor == "Ultralight" ~ 5,
    TRUE ~ NA_real_  # Catch-all for unmatched entries
  ))

table(smokehis$Menthol)
#Coding - Non-Menthol=1, Menthol=2
smokehis <- smokehis %>%
  mutate(MentholBRc = dplyr::case_when(
    Menthol %in% c("Non-Menthol", "Non-menthol") ~ 0,
    Menthol == "Menthol" ~ 1,
    TRUE ~ NA_real_  
  ))

rm(smokehisf1, smokehisf2, smokehisf3, smokehisf4, smokehisf5)
```

**Quiting**
```{r}
#Quit Attempts between Baseline and Follow-up
#Aim(s): 
  #Reduce dimensionality, while retaining as much information as possible
  #Collapse QuA1, QuA4, and QuA5 into QuA4 (number of quit attempts since baseline)
  #Collapse QuA2 and QuA3 into QuA3 (days currently abstinent)
#Note:
  #QuA5 - very difficult to clean (e.g., estimate)
  #See if the following IDs are miss-classified by the model: LV093

quitf <- cessationP3[ , c("ID", "QuA1", "QuA2", "QuA3", "QuA4", "QuA5", "QuA6_1", "QuA6_2", "QuA6_3", "QuA6_4", "QuA6_5", "QuA6_6", "QuA6_7", "QuA6_8", "QuA6_9", "QuA6_10", "QuA6_11", "QuA6_11_TEXT")]

quitff1 <- quitf[which(quitf$QuA1==0), ] #selecting cases in quitf for which QuA1=0 (no) to see cases where number of quit attempts since baseline is 0, but QuA4 and QuA5 > 0 (ID: LV093)
#Coding - Do not change. Response pattern does not make sense and participant may not have been clear on skip logic. If responses were not thoughtfully entered, best not to train the model using this case as an example of quit attempt.

quitff2 <- quitf[which(quitf$QuA1==1), ] #selecting cases in quitf for which QuA1=1 (yes)

quitff3 <- quitf[which(quitf$QuA1==1 & is.na(quitf$QuA4)), ] #selecting cases in quitf for which QuA1=1 (yes), but number of attempts is not listed (IDs: LV001, LV025, LV181, LV241)
#Coding - Can confidently assume at least one quit attempt was made, given QuA1 was endorsed.

quitf$QuA4Rc <- quitf$QuA4
quitf$QuA4Rc[quitf$ID == "LV001" & is.na(quitf$QuA4)] <- 1
quitf$QuA4Rc[quitf$ID == "LV025" & is.na(quitf$QuA4)] <- 1
quitf$QuA4Rc[quitf$ID == "LV181" & is.na(quitf$QuA4)] <- 1
quitf$QuA4Rc[quitf$ID == "LV241" & is.na(quitf$QuA4)] <- 1

quitff4 <- quitf[which(quitf$QuA1==1 & quitf$QuA4==0), ] #selecting cases in quitf for which QuA1=1 (yes), but number of attempts is 0 (IDs: LV050, LV173)
#Coding - Can confidently assume at least one quit attempt was made, given QuA1 was endorsed.

quitf$QuA4Rc[quitf$ID == "LV050" & quitf$QuA4==0] <- 1
quitf$QuA4Rc[quitf$ID == "LV173" & quitf$QuA4==0] <- 1

quitf$QuA6_1Rc <- quitf$QuA6_1
quitf$QuA6_2Rc <- quitf$QuA6_2
quitf$QuA6_3Rc <- quitf$QuA6_3
quitf$QuA6_4Rc <- quitf$QuA6_4
quitf$QuA6_5Rc <- quitf$QuA6_5
quitf$QuA6_6Rc <- quitf$QuA6_6
quitf$QuA6_7Rc <- quitf$QuA6_7
quitf$QuA6_8Rc <- quitf$QuA6_8
quitf$QuA6_9Rc <- quitf$QuA6_9
quitf$QuA6_10Rc <- quitf$QuA6_10
quitf$QuA6_11Rc <- quitf$QuA6_11

quitf <- quitf %>% #replace NA with 0 on multiple columns by name
  mutate(QuA4Rc = coalesce(QuA4Rc, 0), 
         QuA6_1Rc = coalesce(QuA6_1Rc, 0),
         QuA6_2Rc = coalesce(QuA6_2Rc, 0),
         QuA6_3Rc = coalesce(QuA6_3Rc, 0),
         QuA6_4Rc = coalesce(QuA6_4Rc, 0),
         QuA6_5Rc = coalesce(QuA6_5Rc, 0),
         QuA6_6Rc = coalesce(QuA6_6Rc, 0),
         QuA6_7Rc = coalesce(QuA6_7Rc, 0),
         QuA6_8Rc = coalesce(QuA6_8Rc, 0),
         QuA6_9Rc = coalesce(QuA6_9Rc, 0),
         QuA6_10Rc = coalesce(QuA6_10Rc, 0),
         QuA6_11Rc = coalesce(QuA6_11Rc, 0))

quitf$QuA6Tot <- quitf$QuA6_1Rc + quitf$QuA6_2Rc + quitf$QuA6_3Rc + quitf$QuA6_4Rc + quitf$QuA6_5Rc + quitf$QuA6_6Rc + quitf$QuA6_7Rc + quitf$QuA6_8Rc + quitf$QuA6_9Rc + quitf$QuA6_10Rc + quitf$QuA6_11Rc

quitff5 <- quitf[which(quitf$QuA1==1), ]
#Coding - If QuA6Tot=0 (no method listed), code QuA6_11Rc (other) as "1" as best guess

quitf$QuA6_11Rc[quitf$ID == "LV001" & quitf$QuA6Tot==0] <- 1
quitf$QuA6_11Rc[quitf$ID == "LV185" & quitf$QuA6Tot==0] <- 1

quitff6 <- quitf[which(is.na(quitf$QuA1) & quitf$QuA2==0), ]

rm(quitff1, quitff2, quitff3, quitff4, quitff5, quitff6)
```

**Job**
```{r}
job <- cessationP3[ , c("ID", "Job_1", "Job_2", "Job_3", "Job_4", "Job_5", "Job_6", "Job_7", "Job_8", "Job_9")] #select job variables into new data frame

job$Job_1Rc <- job$Job_1
job$Job_2Rc <- job$Job_2
job$Job_3Rc <- job$Job_3
job$Job_4Rc <- job$Job_4
job$Job_5Rc <- job$Job_5
job$Job_6Rc <- job$Job_6
job$Job_7Rc <- job$Job_7
job$Job_8Rc <- job$Job_8
job$Job_9Rc <- job$Job_9

job <- job %>% #replace NA with 0 on multiple columns by name
  mutate(Job_1Rc = coalesce(Job_1Rc, 0), 
         Job_2Rc = coalesce(Job_2Rc, 0),
         Job_3Rc = coalesce(Job_3Rc, 0),
         Job_4Rc = coalesce(Job_4Rc, 0),
         Job_5Rc = coalesce(Job_5Rc, 0),
         Job_6Rc = coalesce(Job_6Rc, 0),
         Job_7Rc = coalesce(Job_7Rc, 0),
         Job_8Rc = coalesce(Job_8Rc, 0),
         Job_9Rc = coalesce(Job_9Rc, 0))

job$JobTot <- job$Job_1Rc + job$Job_2Rc + job$Job_3Rc + job$Job_4Rc + job$Job_5Rc + job$Job_6Rc + job$Job_7Rc + job$Job_8Rc + job$Job_9Rc

jobf1 <- job[which(job$JobTot==0 | job$JobTot==2 | job$JobTot==3), ]
jobf2 <- job[which(job$JobTot==1), ]
#Coding - Collapse into categories of Full-Time Employed = 3, Part-time Employed = 2, Unemployed = 1
  #If missing, code as Unknown (JobRc=0) 
  #If Disabled (Job_5=1), Leave of Absence (Job_6=1), Unemployed (Job_7=1), Retired (Job_8=1), or Homemaker (Job_9=1) is endorsed as only response, code as Unemployed (JobRc=1)
  #If Part-time for pay (Job_3=1) or Part-Time student (Job_4=1) are endorsed as only response, code as Part-Time Employed (JobRc=2)
  #If Full-time (35 hours or more per week) for pay (Job_1=1) or Full-Time student (Job_2=1) is endorsed as only response or with other responses, code as Full-Time Employed (JobRc=3)
  #If Disabled (Job_5=1) and Leave of Absence (Job_6=1), Unemployed (Job_7=1), Retired (Job_8=1), or Homemaker (Job_9=1) is endorsed, code as Unemployed (JobRc=1)
  #If Disabled (Job_5=1) and Full-time (35 hours or more per week) for pay (Job_1=1) or Full-time student (Job_2=1) is endorsed, code as Full Time Employed (JobRc=3)
  #If Disabled (Job_5=1) and Part-time for pay (Job_3=1), or Part-time student (Job_4=1) is endorsed, code as Part Time Employed (JobRc=2)

job$JobRc <- NA

job$JobRc[is.na(job$Job_1) & is.na(job$Job_2) & is.na(job$Job_3) & is.na(job$Job_4) & (job$Job_5==1 | job$Job_6==1 | job$Job_7==1 | job$Job_8==1 | job$Job_9==1)] <- 1

job$JobRc[is.na(job$Job_1) & is.na(job$Job_2) & (job$Job_3 == 1 | job$Job_4 == 1) & is.na(job$Job_5) & is.na(job$Job_6) & is.na(job$Job_7) & is.na(job$Job_8) & is.na(job$Job_9)] <- 2

job$JobRc[is.na(job$Job_1) & is.na(job$Job_2) & (job$Job_3 == 1 | job$Job_4 == 1) & job$Job_5 == 1 & is.na(job$Job_6) & is.na(job$Job_7) & is.na(job$Job_8) & is.na(job$Job_9)] <- 2

job$JobRc[(job$Job_1 == 1 | job$Job_2 == 1) & is.na(job$Job_3) & is.na(job$Job_4) & is.na(job$Job_5) & is.na(job$Job_6) & is.na(job$Job_7) & is.na(job$Job_8) & is.na(job$Job_9)] <- 3

job$JobRc[(job$Job_1 == 1 | job$Job_2 == 1) & is.na(job$Job_3) & is.na(job$Job_4) & job$Job_5 == 1 & is.na(job$Job_6) & is.na(job$Job_7) & is.na(job$Job_8) & is.na(job$Job_9)] <- 3

jobf3 <- job[which(is.na(job$JobRc)), ] #selecting cases in job for which JobRc=NA (IDs: LV004, LV011, LV048, LV069, LV086, LV153, LV170, LV209, LV236, LV246, LV258, LV293, LV299
#Coding - Handcode
  #If Full-Time is endorsed, regardless of other responses code as Full-Time Employed (JobRc=3)
  #If not Full-Time, but Part-Time is endorsed, regardless of other responses code as Part-Time Employed (JobRc=2)
  #LV256 = Missing(NA)

job$JobRc[job$ID == "LV004"] <- 3
job$JobRc[job$ID == "LV069"] <- 2
job$JobRc[job$ID == "LV086"] <- 3

job$JobRc[job$ID == "LV153"] <- 2
job$JobRc[job$ID == "LV209"] <- 3
job$JobRc[job$ID == "LV236"] <- 2
job$JobRc[job$ID == "LV246"] <- 2

job$JobRc[job$ID == "LV258"] <- 3
job$JobRc[job$ID == "LV293"] <- 2

table(job$JobRc)
table(job$Job_1)
table(job$Job_2)

rm(jobf1, jobf2, jobf3)
```

**Stages of Change**
```{r}
#Aim(s): 
  #Reduce dimensionality, while retaining as much information as possible
  #Clean fill-in answers
#Note:
  #SoC3-4, SoC6-7 - very difficult to clean (e.g., estimate)
  #Consider using ContemplationStatus question instead of SoC1-7, since quit attempt history may be sufficient.

stagechange <- cessationP3[ , c("ID", "SoC1", "SoC2", "SoC3", "SoC4", "SoC5", "SoC6", "SoC7")] 
#Coding - No=0, Yes=1 

stagechange <- stagechange %>%
  mutate(
    SoC1BRc = ifelse(SoC1 == 1, 1, ifelse(SoC1 == 2, 0, NA)),
    SoC2BRc = ifelse(SoC2 == 1, 1, ifelse(SoC2 == 2, 0, NA)),
    SoC5BRc = ifelse(SoC5 == 1, 1, ifelse(SoC5 == 2, 0, NA))
  )
```

**E-cigarettes**
```{r}
ecigs <- cessationP3[ , c("ID", "Eciglifetime","Ecigpastmonth", "Brand_ecig", "Menthol_ecig", "Type_ecig")] #select e-cigarette baseline variables

ecigsf1 <- ecigs[which(ecigs$Eciglifetime==0 & is.na(ecigs$Ecigpastmonth)), ]
ecigsf2 <- ecigs[which(ecigs$Eciglifetime==0 & ecigs$Ecigpastmonth==1), ]
ecigsf3 <- ecigs[which(ecigs$Eciglifetime==1 & is.na(ecigs$Ecigpastmonth)), ]

#Coding - Yes=1, No=0, Missing(NA)=0
ecigs <- ecigs %>%
  mutate(
    EcigpastmonthBRc = case_when(
      Ecigpastmonth == 1 ~ 1,
      Eciglifetime == 0 & is.na(Ecigpastmonth) ~ 0,
      Ecigpastmonth == 0 ~ 0,
      TRUE ~ NA_real_
    )
  )

rm(ecigsf1, ecigsf2, ecigsf3)
```

**SSEQ**
```{r}
SSEQb <- cessationP3[, c("ID", "SSEQ1", "SSEQ2_1", "SSEQ3", "SSEQ4_1", "SSEQ5", "SSEQ6_1", "SSEQ7", "SSEQ8_1", "SSEQ9", "SSEQ10_1", "SSEQ11", "SSEQ12_1", "SSEQ13", "SSEQ14_1", "SSEQ15", "SSEQ16_1", "SSEQ17", "SSEQ18_1", "SSEQ19", "SSEQ20_1", "SSEQ21", "SSEQ22_1", "SSEQ23", "SSEQ24_1", "SSEQ25", "SSEQ26_1", "SSEQ27", "SSEQ28_1", "SSEQ29", "SSEQ30_1", "SSEQ31", "SSEQ32_1", "SSEQ33", "SSEQ34_1")]
sapply(SSEQb, function(x) sum(is.na(x))) #total missing values ("NA") in each column
missingtable <- table(SSEQb$SSEQ1, SSEQb$SSEQ2_1, exclude=NULL) #SSEQ1 will be rows, SSEQ2_1 will be columns
missingtable
#Coding - Example, refer to: SSEQ1 and SSEQ2_1 (coded into the even variable). High Control=4, Low Control=3, No Control=2, Does Not Apply=1
  #If SSEQ1 = 1, split SSEQ2_1 high/low (>=51=4 and <=50, >0=3).
  #If SSEQ1 = 1 and SSEQ2_1 = 0, SSEQ2Rc = 2.
  #If SSEQ1 = 1 and SSEQ2_1 = NA, SSEQ2Rc = 3. Conservative guess of low.
  #If SSEQ1 = 2, SSEQ2Rc=2. Also If SSEQ1 = 2 and SSEQ2_1 is endorsed, SSEQ2Rc=2.
  #If SSEQ1 = 3, SSEQ2Rc=1. If SSEQ1 = 3 and SSEQ2_1 is endorsed, SSEQ2Rc=1.
  #If SSEQ1 = NA and SSEQ2_1 is endorsed, split SSEQ2_1 high/low (>=51=4 and <=50, >0=3).
  #If SSEQ1 = NA and SSEQ2_1 = NA, SSEQ2Rc = NA.

for (i in seq(1, 33, 2)) { #loop for SSEQb
  j <- i + 1 #j is i (number beginning with 1) +1
  base <- paste("SSEQ", i, sep = "") #creates label "base" (SSEQ with 1)
  followup <- paste("SSEQ", j, sep = "") #creates label "follow-up" (SSEQ starting with 2)
  followup <- paste(followup, "1", sep = "_") #creates label "follow-up" (SSEQ starting with 2 plus underscore 1)
  output <- paste("SSEQ", j, sep="") #creates label "output" (SSEQ starting with 2)
  output <- paste(output, "BRc", sep = "") #creates label "output" (SSEQ starting with 2 plus Rc)

  SSEQb[[output]][SSEQb[[base]]==1 & SSEQb[[followup]]>=51] <- 4
  SSEQb[[output]][SSEQb[[base]]==1 & SSEQb[[followup]]<=50 & SSEQb[[followup]]>0] <- 3
  SSEQb[[output]][SSEQb[[base]]==1 & is.na(SSEQb[[followup]])] <- 3
  SSEQb[[output]][SSEQb[[base]]==1 & SSEQb[[followup]]==0] <- 2
  SSEQb[[output]][SSEQb[[base]]==2] <- 2
  SSEQb[[output]][SSEQb[[base]]==3] <- 1
  SSEQb[[output]][is.na(SSEQb[[base]]) & SSEQb[[followup]]>=51] <- 4
  SSEQb[[output]][is.na(SSEQb[[base]]) & SSEQb[[followup]]<=50 & SSEQb[[followup]]>0] <- 3
}

sapply(SSEQb, function(x) sum(is.na(x))) #total missing values ("NA") in each column
missingcheck <- SSEQb[, c("ID", "SSEQ1", "SSEQ2_1", "SSEQ2BRc", "SSEQ7", "SSEQ8_1", "SSEQ8BRc")]

rm(base, followup, i, j, output, missingtable, missingcheck)
```

**Coded**
```{r}
codedb <- cessationP3[, c("ID", "PRBQOE_Coded", "SCQOE_Coded", "SGO_Coded")]
#Coding - Perceived benefit=1, Perceived risk=2
codedb$PRBQOEBRc <- codedb$PRBQOE_Coded

#Coding - Positive reinforcement=1, Negative consequence=2, Negative reinforcement=3, Weight control=4
codedb$SCQOEBRc <- codedb$SCQOE_Coded

#Coding (SGO_Coded) - Positive reinforcement=1, Negative consequence=2, Negative reinforcement=3, Neutral action=4, Other/don't know=5
codedb$SGOOEBRc <- codedb$SGO_Coded
```

## Merge New Variables (Part 1)
```{r}
head(smokehis)
smokehism <- subset(smokehis, select = -c(Cig1, Cig2, Cig3, Cig4, Cig5, Cig6, Cig7, Cig8, Cig9, Cig10, Cig11, Brand, Flavor, Menthol, Smokedinpast24hours, hourssincelastcig, Previousquitattempt, COReading, cigsinpast24hours, Avg.cigsday)) 

head(quitf)
quitfm <- subset(quitf, select = -c(QuA1, QuA2, QuA3, QuA4, QuA5, QuA6_1, QuA6_2, QuA6_3, QuA6_4, QuA6_5, QuA6_6, QuA6_7, QuA6_8, QuA6_9, QuA6_10, QuA6_11, QuA6_11_TEXT, QuA6Tot))

head(job)
jobm <- subset(job, select = -c(Job_1, Job_2, Job_3, Job_4, Job_5, Job_6, Job_7, Job_8, Job_9, Job_1Rc, Job_2Rc, Job_3Rc, Job_4Rc, Job_5Rc, Job_6Rc, Job_7Rc, Job_8Rc, Job_9Rc, JobTot)) 

head(stagechange)
stagechangem <- subset(stagechange, select = -c(SoC1, SoC2, SoC3, SoC4, SoC5, SoC6, SoC7))

head(ecigs)
ecigsm <- subset(ecigs, select = -c(Eciglifetime, Ecigpastmonth, Brand_ecig, Menthol_ecig, Type_ecig)) 

head(SSEQb)
SSEQbm <- subset(SSEQb, select = -c(SSEQ1, SSEQ2_1, SSEQ3, SSEQ4_1, SSEQ5, SSEQ6_1, SSEQ7, SSEQ8_1, SSEQ9, SSEQ10_1, SSEQ11, SSEQ12_1, SSEQ13, SSEQ14_1, SSEQ15, SSEQ16_1, SSEQ17, SSEQ18_1, SSEQ19, SSEQ20_1, SSEQ21, SSEQ22_1, SSEQ23, SSEQ24_1, SSEQ25, SSEQ26_1, SSEQ27, SSEQ28_1, SSEQ29, SSEQ30_1, SSEQ31, SSEQ32_1, SSEQ33, SSEQ34_1))

head(codedb)
codedbm <- subset(codedb, select = -c(PRBQOE_Coded, SCQOE_Coded, SGO_Coded))

list1 <- list(codedbm, SSEQbm, ecigsm, stagechangem, jobm, quitfm, smokehism, cessationP3)
merge1 <- list1 %>% reduce(full_join, by='ID')
merge1 <- merge1[,order(colnames(merge1))] #sort columns alphabetically
merge2 <- merge1[order(merge1$ID, decreasing=FALSE),] #sort rows by ID
```

## Remove Variables (Part 2)
```{r}
cessationP4 <- subset(merge2, select = -c(AdminIssuef, Amtmarijuanasmoked, Brand, Brand_ecig, Cig8, Cig9, Cig10, Cig11, CSCLSTotal, Dropout, Ecigpastmonth, Flavor, FTND_Total, hourssincelastcig, Image1, Image2, Image3, ImageResponse_coded, Job_1, Job_2, Job_3, Job_4, Job_5, Job_6, Job_7, Job_8, Job_9, looming, Menthol, Menthol_ecig, narrator, PRBQ_Ben, PRBQ_OE, PRBQ_Risk, PRBQOE_Coded, Previousquitattempt, QuA2, QuA3, QuA4, QuA4Rc, QuA5, QuA6_1, QuA6_10, QuA6_10Rc, QuA6_11, QuA6_11_TEXT, QuA6_11Rc, QuA6_1Rc, QuA6_2, QuA6_2Rc, QuA6_3, QuA6_3Rc, QuA6_4, QuA6_4Rc, QuA6_5,  QuA6_5Rc, QuA6_6, QuA6_6Rc, QuA6_7, QuA6_7Rc, QuA6_8, QuA6_8Rc, QuA6_9, QuA6_9Rc, QuitfMissRc, SCQ_OE, SCQA_Ben, SCQA_Risk, SCQOE_Coded, SGO_Coded, SGO1, Smokedinpast24hours, SoC1, SoC2, SoC3, SoC4, SoC5, SoC6, SoC7, SSEQ_Answered, SSEQ_Total, SSEQ1, SSEQ2_1, SSEQ3, SSEQ4_1, SSEQ5, SSEQ6_1, SSEQ7, SSEQ8_1, SSEQ9, SSEQ10_1, SSEQ11, SSEQ12_1, SSEQ13, SSEQ14_1, SSEQ15, SSEQ16_1, SSEQ17, SSEQ18_1, SSEQ19, SSEQ20_1, SSEQ21, SSEQ22_1, SSEQ23, SSEQ24_1, SSEQ25, SSEQ26_1, SSEQ27, SSEQ28_1, SSEQ29, SSEQ30_1, SSEQ31, SSEQ32_1, SSEQ33, SSEQ34_1, SSSVTotalScore, SysMissf, Type_ecig, VVIQOpen, VVIQClosed))
write.csv(cessationP4, "~/Desktop/LV/cessationP4.csv", row.names=FALSE)

cessationTotal <- merge2 %>% 
  select(ID, CSCLSTotal, FTND_Total, PRBQ_Ben, PRBQ_Risk, SCQA_Ben, SCQA_Risk, SSSVTotalScore, VVIQClosed, VVIQOpen, QuA1)
write.csv(cessationTotal, "~/Desktop/LV/cessationTotal.csv", row.names=FALSE)
                      
rm(cessationP2, codedb, ecigs, job, quitf, smokehis, SSEQb, stagechange, codedbm, ecigsm, jobm, quitfm, smokehism, SSEQbm, stagechangem, list1, merge1, merge2) 
```

## Missing Data (Part 2)
```{r}
#Flag
  #Total: Missing(NA) = 2%
  #Row: Max = 33.06% (96, 154)
  #Column: Max = 18.18 (PRBQOEBRc)
cessationP4[cessationP4 == "NA"] <- NA

vis_miss(cessationP4)
(sum(is.na(cessationP4))/prod(dim(cessationP4)))*100 #percentage of missing data in entire data frame

colmiss <- (colMeans(is.na(cessationP4)))*100 #percentage of missing data by column
colmiss <- as.data.frame(colmiss)
#Coding - If missing >50% by column, exclude variable

mean(complete.cases(cessationP4))*100 #percentage of complete rows
rowmiss <- (rowSums(is.na(cessationP4)))/248*100
rowmiss <- as.data.frame(rowmiss)
#Coding - If missing >50% by row, exclude case

rm(rowmiss, colmiss)
```

## Recode Variables (Part 3)
```{r}
#Coding - QuA1Rc: Yes=1, No=2, 
cessationP4$QuA1Rc <- NA
cessationP4 <- cessationP4 %>%
  mutate(
    QuA1Rc = case_when(
      QuA1 == 1 ~ 1,
      QuA1 == 0 ~ 2,
      TRUE ~ NA_real_
    )
  )

cessationP5 <- subset(cessationP4, select = -c(QuA1))
write.csv(cessationP5, "~/Desktop/LV/cessationP5.csv", row.names=FALSE)

rm(cessationP3) 
```

## Exploratory (Part 1): Univariate

**Data Summary**
```{r}
trainsummary <- summary(cessationP5)
trainsummary <- as.data.frame(trainsummary)
trainsummary <- subset(trainsummary, select = -c(Var1))
trainsummary <- na.omit(trainsummary)
colnames(trainsummary)[1] <- "Variable"

rm(trainsummary)
```

**Categorical Data**

*Features*
```{r}
library(DataExplorer)

plot_intro(cessationP5)
plot_bar(cessationP5)
plot_missing(cessationP5)
```

*Dependent Variable (QuA1)*
```{r}
summary(cessationP5$QuA1Rc)
counts <- table(cessationP5$QuA1Rc)
barplot(counts, main="Quit Attempt",
   xlab="Yes/No")
table(cessationP5$QuA1Rc)
rm(counts)
```

**Continuous Data**

*Age*
```{r}
summary(cessationP5$Age)

hist(cessationP5$Age,
  xlab = "Age",
  main = "Histogram of Age",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Age)$out
outlist <- which(cessationP5$Age %in% c(out))
outlist
boxplot(cessationP5$Age,
  ylab = "Age",
  main = "Boxplot of Ages"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$Age)

rm(out, outlist)
```

*Smoking*
```{r}
#Avg.cigsday
  #Flag: unknown calculation method, remove variable
summary(cessationP5$Avg.cigsday)
hist(cessationP5$Avg.cigsday,
  xlab = "Average CPD",
  main = "Histogram of Average Cigarettes Per Day",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Avg.cigsday)$out
outlist <- which(cessationP5$Avg.cigsday %in% c(out))
outlist
boxplot(cessationP5$Avg.cigsday,
  ylab = "Average Cigarettes per Day",
  main = "Boxplot of Average Cigarettes per Day"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
cor(cessationP5$Avg.cigsday, cessationP5$COReading)
table(cessationP5$Avg.cigsday)
range(cessationP5$Avg.cigsday)

#Cig1
summary(cessationP5$Cig1)
hist(cessationP5$Cig1,
  xlab = "Age of First Cigarette",
  main = "Histogram of Age of First Cigarette",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins

out <- boxplot.stats(cessationP5$Cig1)$out
outlist <- which(cessationP5$Cig1 %in% c(out))
outlist
boxplot(cessationP5$Cig1,
  ylab = "Age of First Cigarette",
  main = "Boxplot of Age of First Cigarette"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
agecigs <- cessationP5[ , c("ID", "Age", "Cig1")] #check that age of first cigarette is younger than study age
agecigs$agediscrep <- 0
agecigs$agediscrep[agecigs$Age < agecigs$Cig1] <- 1
table(agecigs$agediscrep)

#Cig2
summary(cessationP5$Cig2)
hist(cessationP5$Cig2,
  xlab = "Years a Daily Smoker",
  main = "Histogram of Years a Daily Smoker",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Cig2)$out
outlist <- which(cessationP5$Cig2 %in% c(out))
outlist
boxplot(cessationP5$Cig2,
  ylab = "Years a Daily Smoker",
  main = "Boxplot of Years a Daily Smoker"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))

#Cig3
  #Flag: unreliable responses, remove variable
  #Question: What is the minimum number of cigarettes you have smoked on any day in the last 2 months? (Answer in ____ cigarettes)
summary(cessationP5$Cig3)
hist(cessationP5$Cig3,
  xlab = "Minimum CPD",
  main = "Histogram of Minimum Cigarettes Per Day",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Cig3)$out
outlist <- which(cessationP5$Cig3 %in% c(out))
outlist
boxplot(cessationP5$Cig3,
  ylab = "Minimum CPD",
  main = "Boxplot of Minimum CPD"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$Cig3)

#Cig4
  #Flag: unreliable responses, remove variable
  #Question: What is the maximum number of cigarettes you have smoked on any day in the last 2 months? (Answer in ____ cigarettes)
summary(cessationP5$Cig4)
hist(cessationP5$Cig4,
  xlab = "Maximum CPD",
  main = "Histogram of Maximum Cigarettes Per Day",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Cig4)$out
outlist <- which(cessationP5$Cig4 %in% c(out))
outlist
boxplot(cessationP5$Cig4,
  ylab = "Maximum CPD",
  main = "Boxplot of Maximum CPD"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$Cig4)

#Cig5
  #Flag: Keep Cig5 (without adjusting for outliers), keep FTND4, remove Avg.cigsday
  #Question: How many cigarettes do you smoke per day, on average? (Answer in ____ cigarettes)
summary(cessationP5$Cig5)
hist(cessationP5$Cig5,
  xlab = "Average Cigarettes per Day",
  main = "Histogram of Average Cigarettes per Day",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Cig5)$out
outlist <- which(cessationP5$Cig5 %in% c(out))
outlist
boxplot(cessationP5$Cig5,
  ylab = "Average Cigarettes per Day",
  main = "Boxplot of Average Cigarettes per Day"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
cor(cessationP5$Avg.cigsday, cessationP5$Cig5) #r = 0.81
avgcigs <- cessationP5[ , c("ID", "Avg.cigsday", "Cig5", "FTND4")] #compare 3 different questions about average cigarettes per day

avgcigs$Cig5cat <- NA
avgcigs$Cig5cat[avgcigs$Cig5 <=10] <- 0
avgcigs$Cig5cat[avgcigs$Cig5 >=11 & avgcigs$Cig5 <=20] <- 1
avgcigs$Cig5cat[avgcigs$Cig5 >=21 & avgcigs$Cig5 <=30] <- 2
avgcigs$Cig5cat[avgcigs$Cig5 >=31] <- 3
                
avgcigs$avgcat <- NA
avgcigs$avgcat[avgcigs$Avg.cigsday <=10] <- 0
avgcigs$avgcat[avgcigs$Avg.cigsday >=11 & avgcigs$Avg.cigsday <=20] <- 1
avgcigs$avgcat[avgcigs$Avg.cigsday >=21 & avgcigs$Avg.cigsday <=30] <- 2
avgcigs$avgcat[avgcigs$Avg.cigsday >=31] <- 3

cessationP5$Cig5Rc <- cessationP5$Cig5 #replacing missing for Cig5 with median of FTND4 category
cessationP5$Cig5Rc[cessationP5$ID == "LV181" & is.na(cessationP5$Cig5)] <- 5
cessationP5$Cig5Rc[cessationP5$ID == "LV111" & is.na(cessationP5$Cig5)] <- 15

range(cessationP5$Cig5Rc)

#Cig9Rc
summary(cessationP5$Cig9Rc)
hist(cessationP5$Cig9Rc,
  xlab = "Number of Past Quit Attempts",
  main = "Histogram of Number of Past Quit Attempts",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$Cig9Rc)$out
outlist <- which(cessationP5$Cig9Rc %in% c(out))
outlist
boxplot(cessationP5$Cig9Rc,
  ylab = "Number of Past Quit Attempts",
  main = "Boxplot of Number of Past Quit Attempts"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$Cig9Rc)

#cigsinpast24hours
summary(cessationP5$cigsinpast24hours)
hist(cessationP5$cigsinpast24hours,
  xlab = "Number of Cigarettes in the Past 24 Hours",
  main = "Histogram of Number of Cigarettes in the Past 24 Hours",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$cigsinpast24hours)$out
outlist <- which(cessationP5$cigsinpast24hours %in% c(out))
outlist
boxplot(cessationP5$cigsinpast24hours,
  ylab = "Number of Cigarettes in the Past 24 Hours",
  main = "Boxplot of Number of Cigarettes in the Past 24 Hours"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$cigsinpast24hours)

#CL
summary(cessationP5$CL)
hist(cessationP5$CL,
  xlab = "Contemplation Ladder",
  main = "Histogram of Number of Contemplation Ladder",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$CL)$out
outlist <- which(cessationP5$CL %in% c(out))
outlist
boxplot(cessationP5$CL,
  ylab = "Number of Contemplation Ladder",
  main = "Boxplot of Number of Contemplation Ladder"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))

#COReading
summary(cessationP5$COReading)
hist(cessationP5$COReading,
  xlab = "CO Reading",
  main = "Histogram of CO Reading",
  breaks = sqrt(nrow(cessationP5))
) # set number of bins
out <- boxplot.stats(cessationP5$COReading)$out
outlist <- which(cessationP5$COReading %in% c(out))
outlist
boxplot(cessationP5$COReading,
  ylab = "CO Reading",
  main = "Boxplot of CO Reading"
)
mtext(paste("Outliers: ", paste(out, collapse = ", ")))
range(cessationP5$COReading)

rm(out, outlist, avgcigs, agecigs)
```

*CSCLS*
```{r}
CSCLSv <- find_var(cessationP5, pattern = "CSCLS", out = "df")
table(is.na(CSCLSv)) #total missing values ("NA") in dataset 

plot_likert(
  CSCLSv,
  grid.range = c(1.7, 1.9),
  expand.grid = FALSE,
  cat.neutral = 3,
  values = "sum.outside",
  show.prc.sign = TRUE
)

CSCLSv %>%
  pivot_longer(1:6, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 3) +
  labs(x = "Response (Likert 1-5)", y = "")

rm(CSCLSv)
```

*Images*
```{r}
Imagev <- find_var(cessationP5, pattern = "Image", out = "df")
table(is.na(Imagev)) #total missing values ("NA") in dataset 

Imagev %>%
  pivot_longer(1:6, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 3) +
  labs(x = "Response (Likert 1-10)", y = "")

rm(Imagev)
```

*PRBQ*
```{r}
PRBQ1v <- find_var(cessationP5, pattern = "PRBQ1", out = "df")
table(is.na(PRBQ1v)) #total missing values ("NA") in dataset 
PRBQ1v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-7)", y = "")

PRBQ2v <- find_var(cessationP5, pattern = "PRBQ2", out = "df")
table(is.na(PRBQ2v)) #total missing values ("NA") in dataset 
PRBQ2v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-7)", y = "")

PRBQ3v <- find_var(cessationP5, pattern = "PRBQ3", out = "df")
table(is.na(PRBQ3v)) #total missing values ("NA") in dataset 
PRBQ3v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-7)", y = "")

PRBQ4v <- find_var(cessationP5, pattern = "PRBQ4", out = "df")
table(is.na(PRBQ4v)) #total missing values ("NA") in dataset 
PRBQ4v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-7)", y = "")

summary(cessationP5$PRBQ5)
counts5 <- table(cessationP5$PRBQ5)
barplot(counts5, main="PRBQ5",
   xlab="Response (Likert 1-7)")

summary(cessationP5$PRBQ6)
counts6 <- table(cessationP5$PRBQ6)
barplot(counts6, main="PRBQ6",
   xlab="Response (Likert 1-7)")

summary(cessationP5$PRBQ7)
counts7 <- table(cessationP5$PRBQ7)
barplot(counts7, main="PRBQ7",
   xlab="Response (Likert 1-7)")

summary(cessationP5$PRBQ8)
counts8 <- table(cessationP5$PRBQ8)
barplot(counts8, main="PRBQ8",
   xlab="Response (Likert 1-7)")

summary(cessationP5$PRBQ9)
counts9 <- table(cessationP5$PRBQ9)
barplot(counts9, main="PRBQ9",
   xlab="Response (Likert 1-7)")

summary(cessationP5$PRBQOEBRc)
countsOE <- table(cessationP5$PRBQOEBRc)
barplot(countsOE, main="PRBQOE",
   xlab="")

rm(PRBQ1v, PRBQ2v, PRBQ3v, PRBQ4v, counts5, counts6, counts7, counts8, counts9, countsOE)
```

*SCQ*
```{r}
SCQ1v <- find_var(cessationP5, pattern = "SCQ1", out = "df")
table(is.na(SCQ1v)) #total missing values ("NA") in dataset 
SCQ1v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-9)", y = "")

SCQ2v <- find_var(cessationP5, pattern = "SCQ2", out = "df")
table(is.na(SCQ2v)) #total missing values ("NA") in dataset 
SCQ2v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-9)", y = "")

SCQ3v <- find_var(cessationP5, pattern = "SCQ3", out = "df")
table(is.na(SCQ3v)) #total missing values ("NA") in dataset 
SCQ3v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-9)", y = "")

SCQ4v <- find_var(cessationP5, pattern = "SCQ4", out = "df")
table(is.na(SCQ4v)) #total missing values ("NA") in dataset 
SCQ4v %>%
  pivot_longer(1:11, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-9)", y = "")

SCQ5v <- find_var(cessationP5, pattern = "SCQ5", out = "df")
table(is.na(SCQ5v)) #total missing values ("NA") in dataset 
SCQ5v %>%
  pivot_longer(1:7, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-9)", y = "")

summary(cessationP5$SCQ6)
counts6 <- table(cessationP5$SCQ6)
barplot(counts6, main="SCQ6",
   xlab="Response (Likert 1-9)")

summary(cessationP5$SCQ7)
counts7 <- table(cessationP5$SCQ7)
barplot(counts7, main="SCQ7",
   xlab="Response (Likert 1-9)")

summary(cessationP5$SCQ8)
counts8 <- table(cessationP5$SCQ8)
barplot(counts8, main="SCQ8",
   xlab="Response (Likert 1-9)")

summary(cessationP5$SCQ9)
counts9 <- table(cessationP5$SCQ9)
barplot(counts9, main="SCQ9",
   xlab="Response (Likert 1-9)")

summary(cessationP5$SCQOEBRc)
countsOE <- table(cessationP5$SCQOEBRc)
barplot(countsOE, main="SCQOE",
   xlab="")

rm(SCQ1v, SCQ2v, SCQ3v, SCQ4v, SCQ5v, counts6, counts7, counts8, counts9, countsOE)
```

*VVIQ*
```{r}
VVIQ1v <- find_var(cessationP5, pattern = "VVIQ1", out = "df")
table(is.na(VVIQ1v)) #total missing values ("NA") in dataset 

VVIQ1v %>%
  pivot_longer(1:16, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 4) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ2v <- find_var(cessationP5, pattern = "VVIQ2", out = "df")
table(is.na(VVIQ2v)) #total missing values ("NA") in dataset 
VVIQ2v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ3v <- find_var(cessationP5, pattern = "VVIQ3", out = "df")
table(is.na(VVIQ3v)) #total missing values ("NA") in dataset 
VVIQ3v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ4v <- find_var(cessationP5, pattern = "VVIQ4", out = "df")
table(is.na(VVIQ4v)) #total missing values ("NA") in dataset 
VVIQ4v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ5v <- find_var(cessationP5, pattern = "VVIQ5", out = "df")
table(is.na(VVIQ5v)) #total missing values ("NA") in dataset 
VVIQ5v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ6v <- find_var(cessationP5, pattern = "VVIQ6", out = "df")
table(is.na(VVIQ6v)) #total missing values ("NA") in dataset 
VVIQ6v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ7v <- find_var(cessationP5, pattern = "VVIQ7", out = "df")
table(is.na(VVIQ7v)) #total missing values ("NA") in dataset 
VVIQ7v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ8v <- find_var(cessationP5, pattern = "VVIQ8", out = "df")
table(is.na(VVIQ8v)) #total missing values ("NA") in dataset 
VVIQ8v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

VVIQ9v <- find_var(cessationP5, pattern = "VVIQ9", out = "df")
table(is.na(VVIQ9v)) #total missing values ("NA") in dataset 
VVIQ9v %>%
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
geom_bar() +
  facet_wrap(vars(question), ncol = 1) +
  labs(x = "Response (Likert 1-5)", y = "")

rm(VVIQ1v, VVIQ2v, VVIQ3v, VVIQ4v, VVIQ5v, VVIQ6v, VVIQ7v, VVIQ8v, VVIQ9v)
```

*SGO*
```{r}
summary(cessationP5$SGO2)
counts1 <- table(cessationP5$SGO2)
barplot(counts1, main="What is the likelihood of this occurring?",
   xlab="")

summary(cessationP5$SGO2)
counts2 <- table(cessationP5$SGO2)
barplot(counts2, main="How do you view this expectation?",
   xlab="")

summary(cessationP5$SGOOEBRc)
countsOE <- table(cessationP5$SGOOEBRc)
barplot(countsOE, main="SGOOE",
   xlab="")

rm(counts1, counts2, countsOE)
```

*VAS*
```{r}
summary(cessationP5$VAS)
table(cessationP5$VAS, useNA = 'always')
```

*Frequencies*
```{r}
freq1 <- table(cessationP5$COReading, useNA = 'always')
freq1 <- as.data.frame(freq1)
colnames(freq1)[1] <- "COReading"
freq2 <- table(cessationP5$Age, useNA = 'always')
freq2 <- as.data.frame(freq2)
colnames(freq2)[1] <- "Age"
freq3 <- table(cessationP5$Avg.cigsday, useNA = 'always')
freq3 <- as.data.frame(freq3)
colnames(freq3)[1] <- "Avg.cigsday"
freq4 <- table(cessationP5$Cig1, useNA = 'always')
freq4 <- as.data.frame(freq4)
colnames(freq4)[1] <- "Cig1"
freq5 <- table(cessationP5$Cig3, useNA = 'always')
freq5 <- as.data.frame(freq5)
colnames(freq5)[1] <- "Cig3"
freq6 <- table(cessationP5$Cig4, useNA = 'always')
freq6 <- as.data.frame(freq6)
colnames(freq6)[1] <- "Cig4"
freq7 <- table(cessationP5$Cig5Rc, useNA = 'always')
freq7 <- as.data.frame(freq7)
colnames(freq7)[1] <- "Cig5Rc"
freq8 <- table(cessationP5$Cig9Rc, useNA = 'always')
freq8 <- as.data.frame(freq8)
colnames(freq8)[1] <- "Cig9Rc"
freq9 <- table(cessationP5$cigsinpast24hours, useNA = 'always')
freq9 <- as.data.frame(freq9)
colnames(freq9)[1] <- "cigsinpast24hours"

cor(cessationP5$cigsinpast24hours, cessationP5$COReading) #r = 0.29

rm(freq1, freq2, freq3, freq4, freq5, freq6, freq7, freq8, freq9)
```

## Remove Variables (Part 3)
```{r}
cessationP6 <- subset(cessationP5, select = -c(Avg.cigsday, Cig3, Cig4, Cig5, FTND4))
write.csv(cessationP6, "~/Desktop/LV/cessationP6", row.names=FALSE)
cessationP6 <- cessationP6[,order(colnames(cessationP6))] #sort variables alphabetically
cessationP6 <- cessationP6 %>%
    dplyr::select("ID", "QuA1Rc", everything())
cessationP6 <- cessationP6[order(cessationP6$ID, decreasing=FALSE),] #sort rows by ID
rm(cessationP4) 
```

## Missing Data (Part 3)
```{r}
#Flag
  #Total: Missing(NA) = 2.1%
  #Row: Max = 33.6065574% (96, 154)
  #Column: Max = 18.1818182% (PRBQOEBRc)

vis_miss(cessationP6)
(sum(is.na(cessationP6))/prod(dim(cessationP6)))*100 #percentage of missing data in entire data frame
colmiss <- (colMeans(is.na(cessationP6)))*100 #percentage of missing data by column
colmiss <- as.data.frame(colmiss)

mean(complete.cases(cessationP6))*100 #percentage of complete rows
rowmiss <- (rowSums(is.na(cessationP6)))/244*100
rowmiss <- as.data.frame(rowmiss)

rm(colmiss, rowmiss)
```

## Recode Variables (Part 4): Class Imbalance

**Categorical**
```{r}
#Coding - 
  #If few per class, collapse into fewer categories
  #Ignore significant class imbalance for now
  #Missing(NA)= Create "Other" group (variable specific)

cessationP6[cessationP6 == "NA"] <- NA

table(cessationP6$OtherStudies, useNA = "always")
#Coding - 
  #No=0, Yes=1
  #Collapsed categories
cessationP6$OtherStudiesRc <- NA
cessationP6$OtherStudiesRc <- dplyr::recode(cessationP6$OtherStudies, "0" = 0, "1" = 1, "2" = 1, "3" = 1, "4" = 1, "5" = 1) 
table(cessationP6$OtherStudiesRc, useNA = "always")

table(cessationP6$PRBQOEBRc, useNA = "always") #imbalanced: Perceived Benefit=179, Perceived Risk=10, Missing(NA)=42,
#Coding - 
  #Perceived Benefit=0, Other=1
  #Collapsed categories
cessationP6$PRBQOEBRc1 <- NA 
cessationP6$PRBQOEBRc1 <- dplyr::recode(cessationP6$PRBQOEBRc, "1" = 0, "2" = 1) 
table(cessationP6$PRBQOEBRc1, useNA = "always")

table(cessationP6$Ethnicity, useNA = "always") #imbalanced: Hispanic or Latino=5, Not Hispanic or Latino=163, Missing(NA)=17
table(cessationP6$Ethnicity, cessationP6$QuA1Rc, useNA = "always") 
#Coding - 
  #Hispanic or Latino=1, Not Hispanic or Latino=0
  #Missing replaced with mode
cessationP6$EthnicityRc <- NA 
cessationP6$EthnicityRc <- dplyr::recode(cessationP6$Ethnicity, "1" = 1, "2" = 0) 
table(cessationP6$EthnicityRc, useNA = "always")

table(cessationP6$Race, useNA = "always") #imbalanced: American Indian or Alaska Native=8, Black or African American=178, White=38, Asian=4, Native Hawaiian or Other Pacific Islander=1, Missing(NA)=2
#Coding - 
  #Other=0, Black or African American=1
  #Collapsed categories
cessationP6$RaceRc <- NA 
cessationP6$RaceRc <- dplyr::recode(cessationP6$Race, "1" = 0, "2" = 1, "3" = 0, "4" = 0, "5" = 0) 
table(cessationP6$RaceRc, useNA = "always")

table(cessationP5$Education, useNA = "always") #imbalanced: Less than seventh grade=1, Seventh grade to ninth grade=2, Partial high school=14, High school diploma or General Equivalency Degree (GED)=80, Some college (at least one year)=65, Technical school or associates degree=24, College diploma (Bachelors degree)=30, Graduate or professional degree=13, Missing(NA)=1
#Coding - 
  #Less than High School Degree=1, High School Degree=2, Some College=3, Technical/Associates=4, College or Graduate=5, Professional Degree=6
cessationP6$EducationRc <- NA 
cessationP6$EducationRc <- dplyr::recode(cessationP6$Education, "1" = 1, "2" = 1, "3" = 1, "4" = 2, "5" = 3, "6" = 4, "7" = 5, "8" = 6)
table(cessationP6$EducationRc, useNA = "always")

table(cessationP6$Occupation, useNA = "always") #imbalanced: Professional (doctor, therapist, engineer, architect, lawyer, clergy, acct, programmer)=27, Managers and administrators (personnel manager/foreman, manufacturing manager)=38, Craftsmen and kindred workers (mechanic, cabinetmaker, carpenter, electrician, baker, photographer)=29, Clerical and sales workers (secretary, typist, sales-clerk, bartender, buyer, sales representative)=54, Laborers and operatives (janitor, cleaner, stock-handler, maid, bull-dozer operator, fireman, freightman)=40, Student=8, Other (homemaker, never employed)=34, Missing(NA)=1
#Coding - 
  #Other (homemaker, never employed, student)=1, Laborers and operatives (janitor, cleaner, stock-handler, maid, bull-dozer operator, fireman, freightman)=2, Clerical and sales workers (secretary, typist, sales-clerk, bartender, buyer, sales representative)=3, Craftsmen and kindred workers (mechanic, cabinetmaker, carpenter, electrician, baker, photographer)=4, Managers and administrators (personnel manager/foreman, manufacturing manager)=5, Professional (doctor, therapist, engineer, architect, lawyer, clergy, acct, programmer)=6
  #Missing replaced with mode
cessationP6$OccupationRc <- NA 
cessationP6$OccupationRc <- dplyr::recode(cessationP6$Occupation, "1" = 6, "2" = 5, "3" = 4, "4" = 3, "5" = 2, "6" = 1, "7" = 1) 
table(cessationP6$OccupationRc, useNA = "always")

table(cessationP6$Income, useNA = "always") #imbalanced: Less than $10,000=73, $10,000-$19,000=51, $20,000-$29,000=37, $30,000-$39,000=23, $40,000-$49,000=16, $50,000-$74,999=17, $75,000-$99,000=5, $100,000-$149,000=5, $150,000-$199,999=2, $200,000 or more=1, Missing(NA)=1
#Coding - 
  #Less than $10,000=1, $10,000-$19,000=2, $20,000-$29,000=3, $30,000-$39,000=4, $40,000-$49,000=5, $50,000-$74,999=6, $75,000 or more=7,
  #Missing replaced with mode
cessationP6$IncomeRc <- NA 
cessationP6$IncomeRc <- dplyr::recode(cessationP6$Income, "1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7, "8" = 7, "9" = 7, "10" = 7)
table(cessationP6$IncomeRc, useNA = "always")

table(cessationP6$BrandBRc, useNA = "always") 
#Coding - 
  #Other=0, [NAME]=1
cessationP6$BrandBRc1 <- NA 
cessationP6$BrandBRc1 <- dplyr::recode(cessationP6$BrandBRc, "1" = 0, "2" = 0, "3" = 0, "4" = 0, "5" = 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, "10" = 1, "11" = 0, "12" = 0, "13" = 0, "14" = 0, "15" = 0)
table(cessationP6$BrandBRc1, useNA = "always")

table(cessationP6$FlavorBRc, useNA = "always") #imbalanced: Bold=1, Full Flavor=87, Regular=126, Light=16, Ultra Light=1 
#Coding - 
  #Bold/Full Flavor=1, Regular=2, Light/Ultra Light=3 
  #Collapsed categories
cessationP6$FlavorBRc1 <- NA 
cessationP6$FlavorBRc1 <- dplyr::recode(cessationP6$FlavorBRc, "1" = 1, "2" = 1, "3" = 2, "4" = 3, "5" = 3)
table(cessationP6$FlavorBRc1, useNA = "always")
```

## Remove Variables (Part 4)
```{r}
cessationP7 <- subset(cessationP6, select = -c(OtherStudies, PRBQOEBRc, Ethnicity, Race, Education, Occupation, Income, BrandBRc, FlavorBRc))
write.csv(cessationP7, "~/Desktop/LV/cessationP7.csv", row.names=FALSE)
cessationR1 <- cessationP7[,order(colnames(cessationP7))] #sort variables alphabetically
cessationR1 <- cessationR1 %>%
    dplyr::select("ID", "QuA1Rc", everything())

rm(cessationP6)
```

## Missing Data (Part 4)
```{r}
vis_miss(cessationR1)
(sum(is.na(cessationR1))/prod(dim(cessationR1)))*100 #percentage of missing data in entire data frame
colmiss <- (colMeans(is.na(cessationR1)))*100 #percentage of missing data by column
colmiss <- as.data.frame(colmiss)

mean(complete.cases(cessationR1))*100 #percentage of complete rows
rowmiss <- rowSums(is.na(cessationR1))/244*100
rowmiss <- as.data.frame(rowmiss)

sapply(cessationR1, function(x) paste(unique(x), collapse = ", "))

rm(colmiss, rowmiss)
```

## Final Data 
```{r}
write.csv(cessationR1, "~/Desktop/LV/cessationR1.csv", row.names=FALSE)
```
