---
title: 'Smoking Cessation: Logistic Regression'
author: "Zoë E. Laky"
date: '20250603'
output: html_notebook
---

## Installing & Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/LV-main/Final_Data")

required_packages <- c(
  'tidyverse', 'caret', 'visdat', 'moments', 'arsenal', 'ranger',
  'sjPlot', 'sjmisc', 'parameters', 'DataExplorer', 'car', 'naniar', 'recipes', 'kernlab', 'PRROC', 'ggplot2', 'boot',
  'dplyr', 'tidyr')

#install new packages
new_packages <- setdiff(required_packages, rownames(installed.packages()))
if (length(new_packages)) install.packages(new_packages)

#load packages
invisible(lapply(required_packages, require, character.only = TRUE))

#citations
sapply(c("tidyverse", "caret", "sjPlot"), citation)

#working directory
cat("Working Directory:", getwd(), "\n")
sessionInfo()

rm(required_packages, new_packages)
```

## Seed
```{r}
#testing seed 
seed <- 842
set.seed(seed)
rnorm(1) #1.377664
exists(".Random.seed")
.Random.seed
set.seed(NULL)
```

## Loading Data
```{r}
read_clean <- function(file_path) {
  df <- read.csv(file_path, header = TRUE, na.strings = c("", "NA"), stringsAsFactors = TRUE)
  df[df == "NA"] <- NA  
  return(df)}

cessationtrain_p05_final <- read_clean("cessationtrain_p05_final.csv")
cessationtest_p05_final <- read_clean("cessationtest_p05_final.csv")

df_list <- list(
  p05_train  = cessationtrain_p05_final, 
  p05_test   = cessationtest_p05_final)

df_list <- lapply(df_list, function(df) {
  df %>%
    arrange(ID) %>%
    mutate(row_id = row_number())})

cessationtrain_p05_final <- df_list$p05_train
cessationtest_p05_final <- df_list$p05_test

rm(df_list)
```

## Variable Types
```{r}
factor_vars_p05 <- c("QuA1Rc", "Cig7", "ContemplationStatus", "EducationRc", 
                     "SoC2BRc", "SoC5BRc", "SSEQ6BRc", "SSEQ8BRc", "SSS17")

convert_vars <- function(df, factor_vars) {
  df %>%
    mutate(across(-ID, as.numeric)) %>%
    mutate(across(all_of(factor_vars), as.factor))}

cessationtrain_p05_final <- convert_vars(cessationtrain_p05_final, factor_vars_p05)
cessationtest_p05_final  <- convert_vars(cessationtest_p05_final,  factor_vars_p05)

dfs <- c(
  "cessationtrain_p05_final",
  "cessationtest_p05_final")

for (df in dfs) {
  QuA1Rc_df <- get(df)
  if ("QuA1Rc" %in% colnames(QuA1Rc_df)) {
    QuA1Rc_df$QuA1Rc <- factor(QuA1Rc_df$QuA1Rc, levels = c("1", "2"), labels = c("X1", "X2"))}
  assign(df, QuA1Rc_df)}
```

#Manual Folds for Repeated 5-fold Cross-Validation
```{r}
identical(cessationtrain_p05_final$ID, cessationtrain_p05_final$ID) #TRUE

y_p05 <- factor(cessationtrain_p05_final$QuA1Rc)

set.seed(843)
manual_folds_p05 <- createMultiFolds(y = y_p05, k = 5, times = 5)
names(manual_folds_p05)
str(manual_folds_p05[[1]]) 
```

#Log Ridge Recipes
```{r}
log_ridge_recipe_p05 <- recipe(QuA1Rc ~ ., data = cessationtrain_p05_final) %>%
  update_role(ID, new_role = "ID") %>% 
  update_role(row_id, new_role = "ID") %>%
  step_impute_mode(all_nominal_predictors()) %>%  
  step_impute_median(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors(), one_hot = FALSE)
```

#Log Ridge (p < 0.05)
*Training*
```{r}
n_fold <- 5
n_repeat <- 5
n_tune <- 101
n_resample <- n_fold * n_repeat

set.seed(seed)
seeds <- vector("list", length = n_resample + 1)
for (i in seq_len(n_resample)) {
  seeds[[i]] <- sample.int(1e6, n_tune)
}
seeds[[n_resample + 1]] <- sample.int(1e6, 1)

train_ctrl <- trainControl(
  method = "repeatedcv",       
  index = manual_folds_p05,           
  seeds = seeds,               
  classProbs = TRUE,
  summaryFunction = prSummary,
  savePredictions = "all")

set.seed(seed)
log_ridge_p05 <- train(
  log_ridge_recipe_p05,
  data = cessationtrain_p05_final,
  method = "glmnet",
  metric = "F",
  trControl = train_ctrl, 
  tuneGrid = expand.grid(alpha = 0, lambda = c(seq(0.1, 10, by=0.1))),
  family = "binomial")

print(log_ridge_p05$bestTune) #0.01
log_ridge_p05_best_predict <- subset(log_ridge_p05$pred, lambda == log_ridge_p05$bestTune$lambda)
log_ridge_p05_predict <- subset(log_ridge_p05$pred)
log_ridge_p05_resample <- log_ridge_p05$resample
confusionMatrix(log_ridge_p05_best_predict$pred, log_ridge_p05_best_predict$obs)
log_ridge_p05_final_model <- log_ridge_p05$finalModel
print(log_ridge_p05_final_model)
log_ridge_p05_result <- log_ridge_p05$results

#check fold
subset(log_ridge_p05$pred, Resample == "Fold1.Rep1")

#explore coefficients and lambda trade-off
plot(log_ridge_p05_final_model, xvar="lambda",label=TRUE)
plot(log_ridge_p05)

#checking threshold
manual_pred_05 <- ifelse(log_ridge_p05$pred$X1 >= 0.5, "X1", "X2")
mean(manual_pred_05 == log_ridge_p05$pred$pred)  # should be ~1 (or exactly 1)
```

*Fitting*
```{r}
best_alpha  <- log_ridge_p05$bestTune$alpha
best_lambda <- log_ridge_p05$bestTune$lambda

train_ctrl <- trainControl(
  method = 'none',
  classProbs = TRUE,
  summaryFunction = prSummary,
  savePredictions = "all")

set.seed(seed)
log_ridge_final_p05 <- train(
  log_ridge_recipe_p05,
  data = cessationtrain_p05_final,
  method = "glmnet",
  metric = "F",
  trControl = train_ctrl, 
  tuneGrid = expand.grid(
    alpha = best_alpha,      
    lambda = best_lambda),
  family = "binomial")

log_ridge_final_p05_final_model <- log_ridge_final_p05$finalModel
plot(log_ridge_final_p05_final_model, xvar="lambda",label=TRUE)

best_alpha
best_lambda
```

*Testing*
```{r}
log_ridge_p05_train_predict <- predict(log_ridge_final_p05, newdata = cessationtrain_p05_final)
log_ridge_p05_test_predict <- predict(log_ridge_final_p05, newdata = cessationtest_p05_final)

confusionMatrix(log_ridge_p05_train_predict, cessationtrain_p05_final$QuA1Rc)
confusionMatrix(log_ridge_p05_test_predict, cessationtest_p05_final$QuA1Rc)
```

*AUPRC: Train and Test Point Estimates*
```{r}
train_probs <- predict(log_ridge_final_p05, newdata = cessationtrain_p05_final, type = "prob")[, "X1"]
test_probs <- predict(log_ridge_final_p05, newdata = cessationtest_p05_final, type = "prob")[, "X1"]

train_obs <- factor(cessationtrain_p05_final$QuA1Rc, levels = c("X2", "X1"), labels = c("no", "yes"))
test_obs <- factor(cessationtest_p05_final$QuA1Rc, levels = c("X2", "X1"), labels = c("no", "yes"))

#calculate PR curves
pr_train <- pr.curve(scores.class0 = train_probs[train_obs == "yes"],
                     scores.class1 = train_probs[train_obs == "no"],
                     curve = TRUE)

pr_test <- pr.curve(scores.class0 = test_probs[test_obs == "yes"],
                    scores.class1 = test_probs[test_obs == "no"],
                    curve = TRUE)

#graph training and testing PRCs
pr_train_df <- as.data.frame(pr_train$curve)
colnames(pr_train_df) <- c("Recall", "Precision", "Threshold")
pr_train_df$Set <- "Train"
pr_test_df <- as.data.frame(pr_test$curve)
colnames(pr_test_df) <- c("Recall", "Precision", "Threshold")
pr_test_df$Set <- "Test"
pr_df <- bind_rows(pr_train_df, pr_test_df)
baseline_train <- mean(pr_train_df$Precision)
baseline_test <- mean(pr_test_df$Precision)
auprc_values <- data.frame(
  Set = c("Train", "Test"),
  AUPRC = c(pr_train$auc.integral, pr_test$auc.integral))

ggplot(pr_df, aes(x = Recall, y = Precision, color = Set)) +
  geom_line(size = 1.2) +
  geom_hline(yintercept = baseline_train, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = baseline_test, linetype = "dashed", color = "red") +
  annotate("text", x = 0.01, y = baseline_train + 0.02, label = "Train Baseline", 
           hjust = 0, color = "blue", size = 4) +
  annotate("text", x = 0.01, y = baseline_test + 0.02, label = "Test Baseline", 
           hjust = 0, color = "red", size = 4) +
  scale_color_manual(values = c("Train" = "blue", "Test" = "red")) +
  labs(
    title = "LogRidge Precision-Recall Curves: Train vs. Test",
    subtitle = paste0("Train AUPRC = ", round(pr_train$auc.integral, 2),
                      ", Test AUPRC = ", round(pr_test$auc.integral, 2)),
    y = "Precision",
    x = "Recall",
    color = "Data Split") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"))

#table of AUCs by resampling
log_ridge_p05_auprc <- log_ridge_p05_best_predict %>%
  mutate(rep = sub(".*\\.(Rep\\d+)", "\\1", Resample))

log_ridge_p05_repeat_auprc <- log_ridge_p05_auprc %>%
  group_by(rep) %>%
  summarise(auprc = pr.curve(
      scores.class0 = X1[obs == "X1"],
      scores.class1 = X1[obs == "X2"],
      curve = FALSE)$auc.integral)
```

*Bootstrap: Testing AUPRC and Brier*
```{r}
#bootstrap
test_labels <- data.frame(
  obs = test_obs, # factor: "no"/"yes"
  prob_yes = test_probs, # numeric: P("yes")
  outcome = as.numeric(test_obs == "yes")) # 1 = yes, 0 = no
strata <- as.numeric(test_labels$obs) # 1 = yes, 0 = no

pos_class <- "yes"
neg_class <- "no"

#AUPRC
auprc_stat <- function(data, indices) {
  d <- data[indices, ]
  pr <- pr.curve(
    scores.class0 = d$prob_yes[d$obs == pos_class],
    scores.class1 = d$prob_yes[d$obs == neg_class],
    curve = FALSE
  )
  pr$auc.integral
} 

set.seed(843)
boot_out <- boot(data = test_labels, statistic = auprc_stat, R = 2000, strata = strata)
boot_out$t0 #0.3670163                     
boot.ci(boot_out, type = c("perc", "bca")) #CI95% ( 0.1835,  0.6214 )

#Brier
brier_stat <- function(data, indices) {
  d <- data[indices, ]
  mean((d$prob_yes - d$outcome)^2)}

set.seed(843)
boot_brier <- boot(data = test_labels, statistic = brier_stat, R = 2000, strata = strata)
boot_brier$t0 #0.1474545                                  
boot.ci(boot_brier, type = c("perc", "bca")) #CI95% ( 0.1150,  0.1772 )
```

*Calibration Curve*
```{r}
n_bins <- 10  
bin_breaks <- seq(0, 1, length.out = n_bins + 1)

test_labels <- test_labels %>%
  mutate(bin = cut(prob_yes, breaks = bin_breaks, include.lowest = TRUE))

calib_df <- test_labels %>%
  group_by(bin, .drop = FALSE) %>%  # keep empty bins so bootstrap columns align
  summarise(
    mean_pred = mean(prob_yes),
    mean_obs = mean(outcome),
    nb = n(),
    .groups = "drop")

print(calib_df)

calib_stat <- function(data, indices) {
  d <- data[indices, ]
  d$bin <- cut(d$prob_yes, breaks = bin_breaks, include.lowest = TRUE)
  sapply(levels(d$bin), function(b) {
    obs_in_bin <- d$outcome[d$bin == b]
    if (length(obs_in_bin) == 0) return(NA_real_)
    mean(obs_in_bin)})}

set.seed(843)
boot_calib <- boot(data = test_labels, statistic = calib_stat, R = 2000, strata = strata)

calib_ci <- data.frame(
  bin   = levels(test_labels$bin),
  lower = apply(boot_calib$t, 2, quantile, probs = 0.025, na.rm = TRUE),
  upper = apply(boot_calib$t, 2, quantile, probs = 0.975, na.rm = TRUE))

calib_df <- calib_df %>%
  mutate(bin = as.character(bin)) %>%
  left_join(calib_ci, by = "bin")

# sanity check: how many bootstrap resamples had a missing/empty bin?
colSums(is.na(boot_calib$t))

# calibration curve with 95% bootstrap CI band
ggplot(calib_df, aes(x = mean_pred, y = mean_obs)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
  geom_point(aes(size = nb), color = "steelblue") +
  geom_line(color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  labs(x = "Mean Predicted Probability", y = "Observed Proportion",
       title = "Calibration Curve with 95% Bootstrap CI",
       size = "Bin N") +
  theme_minimal(base_size = 14)

range(calib_df$mean_pred, na.rm = TRUE) 
```

*Save*
```{r}
save_directory <- "~/Desktop/LV-main/Final_Models/LogReg_Ridge"

file_path <- file.path(save_directory, "log_ridge_p05_resample.csv")
write.csv(log_ridge_p05_resample, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_p05_auroc_importance.csv")
write.csv(log_ridge_p05_auroc_importance, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_p05_auprc_importance.csv")
write.csv(log_ridge_p05_auprc_importance, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_p05_best_predict.csv")
write.csv(log_ridge_p05_best_predict, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_p05_repeat_auprc.csv")
write.csv(log_ridge_p05_repeat_auprc, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_p05_coefficients.csv")
write.csv(log_ridge_p05_coefficients, file = file_path, row.names = FALSE)
```

#Log Ridge Weight Recipes
```{r}
cessationtrain_p05_final$case_weight <- ifelse(cessationtrain_p05_final$QuA1Rc == "X1", 4, 1)
cessationtest_p05_final$case_weight <- ifelse(cessationtest_p05_final$QuA1Rc == "X1", 4, 1)

log_ridge_weight_recipe_p05 <- recipe(QuA1Rc ~ ., data = cessationtrain_p05_final) %>%
  update_role(ID, new_role = "ID") %>% 
  update_role(row_id, new_role = "ID") %>%
  update_role(case_weight, new_role = "case weight") %>%
  step_impute_mode(all_nominal_predictors()) %>%  
  step_impute_median(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors(), one_hot = FALSE)
```

#Log Ridge Weight (p < 0.05)
*Training*
```{r}
n_fold <- 5
n_repeat <- 5
n_tune <- 101
n_resample <- n_fold * n_repeat

set.seed(seed)
seeds <- vector("list", length = n_resample + 1)
for (i in seq_len(n_resample)) {
  seeds[[i]] <- sample.int(1e6, n_tune)
}
seeds[[n_resample + 1]] <- sample.int(1e6, 1)

train_ctrl <- trainControl(
  method = "repeatedcv",       
  index = manual_folds_p05,           
  seeds = seeds,               
  classProbs = TRUE,
  summaryFunction = prSummary,
  savePredictions = "all"
)

set.seed(seed)
log_ridge_weight_p05 <- train(
  log_ridge_weight_recipe_p05,
  data = cessationtrain_p05_final,
  method = "glmnet",
  metric = "F",
  trControl = train_ctrl,
  tuneGrid = expand.grid(
    alpha = 0,
    lambda = c(seq(0.1, 10, by=0.1))),
  family = "binomial",
  weights = cessationtrain_p05_final$case_weight)

print(log_ridge_weight_p05$bestTune)
log_ridge_weight_p05_best_predict <- subset(log_ridge_weight_p05$pred, lambda == log_ridge_weight_p05$bestTune$lambda)
log_ridge_weight_p05_predict <- subset(log_ridge_weight_p05$pred)
log_ridge_weight_p05_resample <- log_ridge_weight_p05$resample
confusionMatrix(log_ridge_weight_p05_best_predict$pred, log_ridge_weight_p05_best_predict$obs)
log_ridge_weight_p05_final_model <- log_ridge_weight_p05$finalModel
print(log_ridge_weight_p05_final_model)
log_ridge_weight_p05_result <- log_ridge_weight_p05$results

#check fold
subset(log_ridge_weight_p05$pred, Resample == "Fold1.Rep1")

#explore coefficients and lambda trade-off
plot(log_ridge_weight_p05_final_model, xvar="lambda",label=TRUE)
plot(log_ridge_weight_p05)
```

*Fitting*
```{r}
best_alpha  <- log_ridge_weight_p05$bestTune$alpha
best_lambda <- log_ridge_weight_p05$bestTune$lambda

train_ctrl <- trainControl(
  method = 'none',
  classProbs = TRUE,
  summaryFunction = prSummary,
  savePredictions = "all")

set.seed(seed)
log_ridge_weight_final_p05 <- train(
  log_ridge_weight_recipe_p05,
  data = cessationtrain_p05_final,
  method = "glmnet",
  metric = "F",
  trControl = train_ctrl,
  tuneGrid = expand.grid(
    alpha = best_alpha,
    lambda = best_lambda),
  family = "binomial",
  weights = cessationtrain_p05_final$case_weight)

log_ridge_weight_final_p05_final_model <- log_ridge_weight_final_p05$finalModel
plot(log_ridge_weight_final_p05_final_model, xvar="lambda",label=TRUE)

best_alpha
best_lambda
```

*Testing*
```{r}
log_ridge_weight_p05_train_predict <- predict(log_ridge_weight_final_p05, newdata = cessationtrain_p05_final)
log_ridge_weight_p05_test_predict <- predict(log_ridge_weight_final_p05, newdata = cessationtest_p05_final)

confusionMatrix(log_ridge_weight_p05_train_predict, cessationtrain_p05_final$QuA1Rc)
confusionMatrix(log_ridge_weight_p05_test_predict, cessationtest_p05_final$QuA1Rc)
```

*AUPRC: Train and Test Point Estimates*
```{r}
train_probs <- predict(log_ridge_weight_final_p05, newdata = cessationtrain_p05_final, type = "prob")[, "X1"]
test_probs <- predict(log_ridge_weight_final_p05, newdata = cessationtest_p05_final, type = "prob")[, "X1"]

train_obs <- factor(cessationtrain_p05_final$QuA1Rc, levels = c("X2", "X1"), labels = c("no", "yes"))
test_obs <- factor(cessationtest_p05_final$QuA1Rc, levels = c("X2", "X1"), labels = c("no", "yes"))

#calculate PR curves
pr_train <- pr.curve(scores.class0 = train_probs[train_obs == "yes"],
                     scores.class1 = train_probs[train_obs == "no"],
                     curve = TRUE)

pr_test <- pr.curve(scores.class0 = test_probs[test_obs == "yes"],
                    scores.class1 = test_probs[test_obs == "no"],
                    curve = TRUE)

#graph training and testing PRCs
pr_train_df <- as.data.frame(pr_train$curve)
colnames(pr_train_df) <- c("Recall", "Precision", "Threshold")
pr_train_df$Set <- "Train"
pr_test_df <- as.data.frame(pr_test$curve)
colnames(pr_test_df) <- c("Recall", "Precision", "Threshold")
pr_test_df$Set <- "Test"
pr_df <- bind_rows(pr_train_df, pr_test_df)
baseline_train <- mean(pr_train_df$Precision)
baseline_test <- mean(pr_test_df$Precision)
auprc_values <- data.frame(
  Set = c("Train", "Test"),
  AUPRC = c(pr_train$auc.integral, pr_test$auc.integral))

ggplot(pr_df, aes(x = Recall, y = Precision, color = Set)) +
  geom_line(size = 1.2) +
  geom_hline(yintercept = baseline_train, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = baseline_test, linetype = "dashed", color = "red") +
  annotate("text", x = 0.01, y = baseline_train + 0.02, label = "Train Baseline", 
           hjust = 0, color = "blue", size = 4) +
  annotate("text", x = 0.01, y = baseline_test + 0.02, label = "Test Baseline", 
           hjust = 0, color = "red", size = 4) +
  scale_color_manual(values = c("Train" = "blue", "Test" = "red")) +
  labs(
    title = "LogRidge Weight Precision-Recall Curves: Train vs. Test",
    subtitle = paste0("Train AUPRC = ", round(pr_train$auc.integral, 2),
                      ", Test AUPRC = ", round(pr_test$auc.integral, 2)),
    y = "Precision",
    x = "Recall",
    color = "Data Split") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"))

#table of AUCs by resampling
log_ridge_weight_p05_auprc <- log_ridge_weight_p05_best_predict %>%
  mutate(rep = sub(".*\\.(Rep\\d+)", "\\1", Resample))

log_ridge_weight_p05_repeat_auprc <- log_ridge_weight_p05_auprc %>%
  group_by(rep) %>%
  summarise(auprc = pr.curve(
      scores.class0 = X1[obs == "X1"],
      scores.class1 = X1[obs == "X2"],
      curve = FALSE)$auc.integral)
```

*Bootstrap: Testing AUPRC and Brier*
```{r}
#bootstrap
test_labels <- data.frame(
  obs = test_obs, # factor: "no"/"yes"
  prob_yes = test_probs, # numeric: P("yes")
  outcome = as.numeric(test_obs == "yes")) # 1 = yes, 0 = no
strata <- as.numeric(test_labels$obs) # 1 = yes, 0 = no

pos_class <- "yes"
neg_class <- "no"

#AUPRC
auprc_stat <- function(data, indices) {
  d <- data[indices, ]
  pr <- pr.curve(
    scores.class0 = d$prob_yes[d$obs == pos_class],
    scores.class1 = d$prob_yes[d$obs == neg_class],
    curve = FALSE
  )
  pr$auc.integral
} 

set.seed(843)
boot_out <- boot(data = test_labels, statistic = auprc_stat, R = 2000, strata = strata)
boot_out$t0 #0.3921733                  
boot.ci(boot_out, type = c("perc", "bca")) #CI95% ( 0.2002,  0.6691 )

#Brier
brier_stat <- function(data, indices) {
  d <- data[indices, ]
  mean((d$prob_yes - d$outcome)^2)}

set.seed(843)
boot_brier <- boot(data = test_labels, statistic = brier_stat, R = 2000, strata = strata)
boot_brier$t0 #0.2216675                                
boot.ci(boot_brier, type = c("perc", "bca")) #CI95% ( 0.2002,  0.6691 )
```

*Calibration Curve*
```{r}
n_bins <- 10  
bin_breaks <- seq(0, 1, length.out = n_bins + 1)

test_labels <- test_labels %>%
  mutate(bin = cut(prob_yes, breaks = bin_breaks, include.lowest = TRUE))

calib_df <- test_labels %>%
  group_by(bin, .drop = FALSE) %>%  # keep empty bins so bootstrap columns align
  summarise(
    mean_pred = mean(prob_yes),
    mean_obs = mean(outcome),
    nb = n(),
    .groups = "drop")

print(calib_df)

calib_stat <- function(data, indices) {
  d <- data[indices, ]
  d$bin <- cut(d$prob_yes, breaks = bin_breaks, include.lowest = TRUE)
  sapply(levels(d$bin), function(b) {
    obs_in_bin <- d$outcome[d$bin == b]
    if (length(obs_in_bin) == 0) return(NA_real_)
    mean(obs_in_bin)})}

set.seed(843)
boot_calib <- boot(data = test_labels, statistic = calib_stat, R = 2000, strata = strata)

calib_ci <- data.frame(
  bin   = levels(test_labels$bin),
  lower = apply(boot_calib$t, 2, quantile, probs = 0.025, na.rm = TRUE),
  upper = apply(boot_calib$t, 2, quantile, probs = 0.975, na.rm = TRUE))

calib_df <- calib_df %>%
  mutate(bin = as.character(bin)) %>%
  left_join(calib_ci, by = "bin")

# sanity check: how many bootstrap resamples had a missing/empty bin?
colSums(is.na(boot_calib$t))

# calibration curve with 95% bootstrap CI band
ggplot(calib_df, aes(x = mean_pred, y = mean_obs)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
  geom_point(aes(size = nb), color = "steelblue") +
  geom_line(color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  labs(x = "Mean Predicted Probability", y = "Observed Proportion",
       title = "Calibration Curve with 95% Bootstrap CI",
       size = "Bin N") +
  theme_minimal(base_size = 14)

range(calib_df$mean_pred, na.rm = TRUE) 
```

*Save*
```{r}
save_directory <- "~/Desktop/LV-main/Final_Models/LogReg_Ridge_Weighted"

file_path <- file.path(save_directory, "log_ridge_weight_p05_resample.csv")
write.csv(log_ridge_weight_p05_resample, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_weight_p05_auroc_importance.csv")
write.csv(log_ridge_weight_p05_auroc_importance, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_weight_p05_auprc_importance.csv")
write.csv(log_ridge_weight_p05_auprc_importance, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_weight_p05_best_predict.csv")
write.csv(log_ridge_weight_p05_best_predict, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_weight_p05_repeat_auprc.csv")
write.csv(log_ridge_weight_p05_repeat_auprc, file = file_path, row.names = FALSE)

file_path <- file.path(save_directory, "log_ridge_weight_p05_coefficients.csv")
write.csv(log_ridge_weight_p05_coefficients, file = file_path, row.names = FALSE)
```

